---
title: "Motivational Transitions in Physical Activity from Youth to Adulthood: Insights from Multigroup SEM and Latent Class Approaches"
date: "2025-10-30"
author: "Student ID: 11484265"
output:
  pdf_document: default
  html_document: default
  output: pdf_document
  fontsize: 12pt
  urlcolor: "blue"
  geometry: margin=2in
---



\pagenumbering{arabic}

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo=FALSE, 
                      eval=TRUE, message=FALSE, 
                      warning=FALSE, fig.align='center',
                      fig.dim=c(5,3),
                      tidy.opts = list(width.cutoff = 60), tidy = TRUE)
                      
library(rmarkdown)
library(tidyverse)
library(tinytex)
library(checkpoint)
library(knitr)
library(kableExtra)
library(manipulateWidget)
library(rmdwc)

library(scales)
library(lavaan)
library(car)
library(ggplot2)
library(nnet)
library(tidyLPA)
library(poLCA)
library(poLCAExtra)

```

\begin{center}
Word count: `r formatC(rmdwc::rmdcount("Paper.Rmd")$words-2871, format = "f", big.mark = ",", digits = 0)`
\end{center}

\newpage 

```{r data, include=FALSE, cache=F}
source(file="data.R")
```

```{r SEM, include=FALSE, cache=F}
source(file="SEM.R")
```

```{r LCA, include=FALSE, cache=F}
source(file="LCA.R")
```

```{r Visualization, include=FALSE, cache=F}
source(file="Visualization.R")
```

```{=latex}
\setcounter{tocdepth}{3}
\tableofcontents
```

\newpage 

# Introduction


Physical inactivity remains a pressing global public health concern, with nearly 1.8 billion individuals failing to meet recommended levels of physical activity (PA) (World Health Organization, 2024), representing a 9% increase over the past two decades (Mitchell, 2019; Strain *et al.*, 2024). Insufficient PA contributes not only to physiological decline but also to a broad spectrum of chronic diseases, including cardiovascular conditions, diabetes, and certain cancers, as well as mental health issues such as depression and anxiety (Anderson & Shivakumar, 2013; Booth *et al.*, 2012; White *et al.*, 2024). Given this multifaceted impact, understanding the drivers of PA behaviour has become a major public health priority. Among these drivers, motivation consistently emerges as one of the most robust predictors of sustained participation (Daley & Duda, 2006; Duncan *et al.*, 2010).

Motivation is central to explaining why individuals initiate, persist, or disengage from PA. Psychological frameworks such as Self-Determination Theory (SDT) and the Theory of Planned Behaviour (TPB) have been foundational in conceptualizing these motivational processes. SDT highlights the role of intrinsic motivation, and the satisfaction of autonomy, competence and relatedness needs in promoting long-term adherence (Deci & Ryan, 2008; Brooks *et al.*, 2017). TPB, in contrast, situates motivation within broader social-cognitive structures, linking behavioural intentions to attitudes, subjective norms, and perceived behavioural control (PBC) (Ajzen, 1991). Yet despite extensive research using these models, most studies have examined age-related motivational differences within either youth or adult populations in isolation. Few longitudinal studies cover the full age span from early adolescence to older adulthood, and even fewer have explicitly compared youths and adults, or examined age differences across configurations of PA motives. Addressing these underexplored areas can inform the design of developmentally tailored interventions and policy initiatives. The following research questions are examined:

1. How does the relationship between perceived motives and PA differ between youths and adults?

2. What is the relationship between age and motivational profiles of PA in youths and adults?

To answer these questions, a multigroup structural equation modelling (SEM) approach was first applied to test whether the relationships between specific motives and PA differ between developmental stages. To capture the multidimensional nature of motivation, latent class analysis (LCA) was then employed to identify distinct motivational profiles within youths and adults separately. Using comparable variables and data structures across both groups allows for a direct assessment of whether similar age-related patterns exist, which will be tested using multinomial logistic regression.

# Literature Review

## Self-Determination Theory

SDT is a popular framework that has gained popularity in the past two decades. Developed by Ryan & Deci (2000), it emphasizes the degree to which behaviour is self-determined. Based on the Basic Needs minitheory, individuals are oriented around three needs, namely autonomy, the belief in having choice, competence, the ability to work effectively, and relatedness, the drive to form bonds with others (Reeve, 2012). These needs in turn drive motivation, which is classified on the autonomy continuum as amotivation, extrinsic, and intrinsic. Within extrinsic motivation, there exist external, introjected, identified, and integrated regulation, with internalization increasing progressively from none to fully intrinsic under the minitheory Organismic Integration Theory within SDT (Standage *et al.*, 2019).

External regulation occurs when actions are guided by outside influences, such as rewards or punishments. Introjected regulation reflects behaviour driven by internal pressures, such as guilt. In identified regulation, individuals willingly engage in an activity because they recognize its personal value or the benefits it provides. Integrated regulation aligns behaviour is completely aligned with one's identity and broader goals (Ryan & Deci, 2017). Intrinsic motivation, by contrast, arises from interest, enjoyment, or inherent satisfaction in the activity itself. While evidence on the relationship between PA and specific constructs is mixed, prior studies indicate that stronger intrinsic motives are positively associated with PA participation, lower dropout rates, and improved physical and psychological wellbeing.(Ng *et al.*, 2012; Teixeira *et al.*, 2012).

Nevertheless, while SDT posits that autonomy-supportive environments foster intrinsic motivation, recent studies suggest that this relationship may not be as straightforward as previously thought. For instance, interventions designed to enhance autonomy support have shown inconsistent effects on long-term PA adherence (Trigueros *et al.*, 2019). While this may be due to individual differences in baseline motivation or contextual factors, past research often overlooked the role of external constraints and structural factors, such as socioeconomic status. These external factors may interact with psychological needs in complex ways that SDT does not fully address.

For clarity, the following terms used in this paper reference to SDT constructs: intrinsic motivation, internalized motivation, extrinsic motivation, externalized motivation, autonomous motivation, controlled regulation, identified regulation, introjected regulation, integrated regulation, amotivation, competence, relatedness.


## Theory of Planned Behaviour

Proposed by Ajzen in 1991, TPB is a widely used framework for predicting and understanding human behaviour, emphasizing the role of intention as the most proximal determinant of action. According to TPB, behavioural intentions are influenced by three key factors: attitudes toward the behaviour, subjective norms, and PBC. Attitudes reflect an individual's positive or negative evaluation of performing the behaviour, subjective norms capture perceived social pressure, and PBC represents the perceived performance ease. These factors interact to shape intention, which in turn predicts behaviour. In the context of PA, TPB has been used to explain variations in PA participation across age groups, demonstrating that stronger intentions, supported by favourable attitudes, positive social norms, and higher PBC, are associated with higher levels of activity (Neipp *et al.*, 2013; Wang & Wang, 2015). 

Like SDT, TPB faces several critiques. Its constructs are typically measured through self-report instruments, which are vulnerable to biases such as social desirability and general inaccuracies. Ajzen (2020) highlighted the complexity of assessing the beliefs relating to TPB components, noting that they do not always correspond perfectly with actual behaviour. In addition, TPB primarily focuses on individual-level factors and may insufficiently consider broader structural influences. Its emphasis on personal perceptions of control can also overlook systemic barriers or facilitators that shape behaviour. Furthermore, the theory's predictive performance varies by gender, and the role of TPB variables in predicting PA intentions is inconsistent. While PBC has been found to explain both intention and behaviour (Ajzen, 1991; Godin & Kok, 1996), other studies identified attitude as the strongest predictor, with subjective norm generally showing significant variation in influence (Wang & Syed Ali, 2025). However, exceptions exist, such as its positive association with behaviour among professional footballers (Moreno-Murcia *et al.*, 2013). More recently, Kim *et al.* (2019) suggested that previous weak effects of subjective norms may have been a result of methodological neglect. Since prior studies often used instruments without consistent standardization (Wang & Syed Ali, 2025), discrepancies in findings may reflect developmental differences or measurement variation. Even when the same instrument is used, motivational items may be interpreted differently across age groups. Nevertheless, there exists consistent evidence for the relationship between PA and TPB variables in general (Hagger *et al.*, 2002), especially PBC and attitude (Wang & Syed Ali, 2025). 

For clarity, the following terms used in this paper reference to TPB constructs: perceived behavioural control, subjective norms, attitude toward behaviour, intention, and behavioural beliefs.


## Age-Related Differences

Intrinsic motivation tends to decline with age regardless of developmental stage (Brunet & Sabiston, 2011; Dishman *et al.*, 2018; de Maio Nascimento *et al.*, 2023), suggesting that the factors driving engagement shift across the lifespan.

Older adults consistently report lower intrinsic motivation and introjected regulation, which corresponds with reduced PA levels compared to younger adults (Brunet & Sabiston, 2011). In contrast, middle-aged adults are more likely to exercise for health, fitness, and psychological well-being, suggesting a shift toward more outcome-oriented motives with age (Brunet & Sabiston, 2011; de Maio Nascimento *et al.*, 2023). Notably, some motivators, such as relaxation and stress relief, remain relatively stable across age groups, indicating that these factors may serve as universal supports for PA adherence and associated psychological benefits (de Maio Nascimento *et al.*, 2023; Vuckovic & Duric, 2024; Kilgour *et al.*, 2005). Contextual factors, including access to facilities and opportunities for participation, interact with PA levels consistently across the adult lifespan (Brunet & Sabiston, 2011; Dishman *et al.*, 2018), highlighting the importance of structural supports alongside motivational processes. These patterns may reflect qualitative shifts in motivational orientation, where self-regulatory and instrumental goals become central. This raises the question of whether lower intrinsic motivation in older adults represents genuine motivational loss or an adaptive reconfiguration.

In youths, intrinsic motivation, enjoyment, and social support emerge as key drivers of habitual PA. Evidence indicates that those who sustain high intrinsic motivation, such as enjoyment, maintain higher activity levels over time (Sebire *et al.*, 2013). Conversely, Dishman *et al.* (2018) found that declines in goals related to social recognition, competence, or appearance can undermine engagement. Past research further link higher motivation to adaptive outcomes such as social support (Heredia-León *et al.*, 2023), which is correlated with higher PA levels (Salvy *et al.*, 2008). These patterns suggest that sustained PA in youths depends on motives that are personally meaningful and enjoyable. Shifts toward externally oriented goals may reduce intrinsic engagement, reflecting a qualitative reorganization of motivation rather than simple decline in PA with age (Shao & Zhou, 2023). Moreover, the relationship between social support and PA may indicate that motivation is embedded within social networks rather than residing solely within the individual, highlighting the importance of social influence.

Although potential multicollinearity (Brunet & Sabiston, 2011), single-region recruitment and missing objective measures (de Maio Nascimento *et al.*, 2023), small subgroups and cultural specificity (Vuckovic & Duric, 2024), and limited demographic focus (Dishman *et al.*, 2018; Kilgour *et al.*, 2024) pose limitations, clear age-related differences emerge. Enjoyment, social motives, and guilt appear particularly important for youths, whereas fitness, relaxation, and psychological well-being motives are more salient in adults. Comparing these patterns allows for distinguishing whether observed age differences reflect developmental shifts in the relevance of particular motives or simply variations in overall motivational intensity. Assessing the differences of these motives provides a more nuanced understanding of how motivation supports PA across the lifespan.

* H1(a). Enjoyment, social, and guilt motives are more strongly associated with physical activity in youths than in adults.

* H1(b). Fitness and relaxation motives are more strongly associated with physical activity in adults than in youths.

* H1(c). Opportunity is similarly associated with physical activity across age groups.

## Motivational Profiles

In adults, Nuss *et al.* (2023) identified four motivational profiles in Canadian adults, showing that combinations of controlled and autonomous motivation synergistically support PA. Ostendorf *et al.* (2021) similarly identified three primary profiles among individuals with overweight or obesity. Longitudinally, those in the high-autonomous profile exhibited the least decline in moderate-to-vigorous PA during transitions from supervised to unsupervised PA, suggesting that intrinsic and identified motives sustain engagement. In youths, Moreno-Murcia *et al.* (2013) identified two primary motivational profiles in physical education students, with membership in the self-determined profile associated with higher PA levels. Heredia-León *et al.* (2023) likewise found that youths in higher motivation profiles demonstrated greater PA intention and enjoyment in PE classes. These findings suggest that the type of motivation may be an important correlate of PA, alongside motivation strength.

However, several limitations constrain interpretation. Existing studies focus on either adults or youths, making cross-stage comparisons speculative. Sample characteristics, such as health status (Ostendorf *et al.*, 2021) or cultural context (Moreno-Murcia *et al.*, 2013), may drive observed profile differences. Methodological variations, including the number and type of motivational items, statistical extraction methods, and criteria for class selection, further complicate cross-study comparisons and obscure developmental trends.

Examining age-related patterns on motivational profiles using comparable variables and analytic approaches across youths and adults could yield a clearer picture of developmental trajectories. This approach examines whether the likelihood of different motivation profiles varies consistently across age groups. Using comparable data structures allows for a direct assessment of these patterns.

* H2(a). With increasing age, adults are more likely to belong to motivational profiles characterized by lower motivation.

* H2(b). With increasing age, youths are more likely to belong to motivational profiles characterized by lower motivation.

Taken together, H2(a) and H2(b) examine whether the age-related changes in profile membership follow a comparable pattern in adults and youths.

# Data and Variables

The analyses draw on large-scale survey data collected by Ipsos on behalf of Sport England (2024, 2025). These datasets were chosen because the youth and adult surveys share a parallel structure, with several items identically worded across both instruments. This allows direct comparison of motive items across life stages. All motive items are self-reported, single-item measures. Appendix A lists the corresponding survey questions. To maintain comparability and avoid skew from small subgroups, only cisgender adults without disabilities were retained. Ethnicity and gender were included as control variables. Due to small group sizes, ethnicity was coded as White British or Other. Observations with missing data or "I don't know" answers on any used item were likewise excluded. Additional descriptive statistics, including percentages of missing values are provided in Appendix C. Weekly minutes of moderate to vigorous PA were collected. For youths, out-of-school PA was used to capture voluntary activity, accounting for differences in curriculum requirements across grade levels. See Appendix B for full list of eligible activities. The following is a list of motive and PA items:

* Enjoyment: finding PA satisfying or pleasurable.

* Social: exercising for fun or connection with friends.

* Fitness: exercising to maintain or improve physical health.

* Guilt: exercising out of obligation or self-pressure.

* Opportunity: having access, time, or suitable conditions to exercise.

* Relaxation: exercising to relieve stress or worry.

* Ability: self-perceived competence and confidence in PA.

* Importance: recognizing the personal value or significance of PA.

* Challenge: persistence and enjoyment in pursuing difficult tasks.

* Minutes: total weekly minutes of moderate-to-vigorous PA.

Although this study did not directly measure constructs from SDT, some of the examined motives can be conceptually mapped onto its continuum of motivational regulation. Enjoyment in PA reflects intrinsic motivation, while exercising to stay fit, to relax, or because one understands its benefits represent forms of identified regulation. Exercising socially for fun with friends can reflect either intrinsic or identified regulation. In contrast, exercising due to guilt is a classic example of introjected regulation. These mappings should be regarded as conceptual approximations rather than definitive classifications.

## Adult Dataset

The adult data derive from the Active Lives Adult Survey (Year 8: 2022-2023), conducted by Ipsos for Sport England with additional support from the Office for Health Improvement and Disparities (OHID). The survey was designed to provide a nationally representative picture of adult participation in PA in England.

Sampling was based on the Postcode Address File (PAF), using a random probability design. Up to two adults (aged 16+) per selected household were invited to participate. A push-to-web approach was employed: two initial mailings invited online completion, followed by a third mailing with a paper questionnaire to increase response rates, and a final reminder letter. Data collection occurred from November 2022 to November 2023, with approximately 500 responses targeted per local authority to minimize seasonal and geographic bias. A total of 173,950 adults completed the survey. Weighting adjustments were applied to correct for differential response probabilities by Sport England. The questionnaire covered a broad range of topics, including PA frequency and duration, sports participation, volunteering, club membership, and motivational factors, along with demographic variables such as age, gender, education, and socio-economic status.


```{r gg.ad.demo1, out.width='32%',fig.show='hold',fig.align='center', fig.cap='Distribution of age, gender, and ethnicity in adults.'}
gg.ad.age
gg.ad.gend
gg.ad.eth
```

Females comprised `r prettyNum(prop.table(table(adult.bi$gender))[[2]]*100, big.mark=",", digits = 2)`%, and White British participants comprised `r prettyNum(prop.table(table(adult.lik$eth))[[1]]*100, big.mark=",", digits = 2)`% of the sample. Given no indication of meaningful differences in motivation and PA during preliminary analyses, the youngest and oldest age groups were collapsed with their neighbours due to small cell sizes at the extremes, forming six age brackets, namely 16-34, 35-44, 45-54, 55-64, 65-74, and 75+ (*Figure 1*). Collapsing these categories improved model stability without obscuring theoretically relevant distinctions. After adjustment, the age distribution remained moderately skewed but acceptable for analysis.

```{r vars.ad, out.width='80%',,fig.show='hold',fig.align='center', fig.cap='Distribution of adult responses for each motive on the Likert scale.'}

gg.vars.ad

```

Each item is rated on a five-point Likert scale ranging from strongly agree (1) to strongly disagree (5). Due to low frequencies in the disagreement categories, negative responses were combined into one to reduce sparsity and minimize distortion. As seen in *Figure 2*, the proportion of adults expressing strong positive agreement declines with age for most motive items. Older adults were less likely than younger adults to strongly endorse beliefs such as feeling capable of engaging in PA, valuing it for the challenge, or pursuing it for enjoyment and fitness. They were also less likely to report strong guilt for not exercising, to perceive sufficient opportunities for activity, or to consider PA an important means of relaxation. 

## Youth Dataset

The youth data come from the Active Lives Children and Young People Survey (Year 6: 2022-2023), conducted by Ipsos on behalf of Sport England, which parallels the adult survey in design and content. It is a large-scale, school-based online survey administered to pupils in Years 1-11, their parents (for Years 1-2 pupils), and teachers. Only pupil responses from year 6-11 (age 11-16) were used in this dissertation due to significant deviations in survey design for other years from the adult version. The sampling strategy was designed to permit analysis at both national and local authority levels. A stratified three-stage sampling process was used: schools were first sampled from the January 2021 school census, then three year groups were randomly selected within each participating school, and finally, one mixed-ability class was chosen per selected year group. Pupils either completed the survey at school under teacher supervision or as homework. To encourage participation, schools received credits for sports equipment and school-level feedback reports. 

```{r gg.ch.demo, out.width='32%',fig.show='hold',fig.align='center', fig.cap='Distribution of age, gender, and ethnicity in youths.'}

gg.ch.age
gg.ch.gend
gg.ch.eth

```

`r prettyNum(prop.table(table(child.lik$gender))[[2]]*100, big.mark=",", digits = 2)`% of the sample are female, and `r prettyNum(prop.table(table(child.lik$eth))[[1]]*100, big.mark=",", digits = 2)`% are White British (*Figure 3*). Although 16-year-olds represent a small proportion of the sample (4.63%), they were not collapsed with the 15-year-olds. The transition from age 15 to 16 marks a distinct developmental and social stage, such as reaching legal thresholds, completing compulsory schooling, and gaining increased autonomy. Retaining this distinction allows examination of possible developmental turning points.

```{r vars.ch,out.width="80%", fig.show='hold',fig.align='center', fig.cap='Distribution of youth responses for each motive on the Likert scale.'}

gg.vars.ch

```

Each item was rated on a four-point Likert scale ranging from strongly agree (1) to strongly disagree (4). As with the adult data, the low frequency, negative categories were collapsed to reduce sparsity. After collapsing, the items are generally balanced (*Figure 4*), except for social, guilt, and relaxation. Response patterns among youths are less consistent than those observed in adults. With increasing age, youths are less likely to report exercising for challenge or social reasons, while a larger proportion indicate exercising due to guilt or for relaxation. Among the oldest youth group, there is a slight but noticeable increase in the proportion who strongly agree that ability and fitness are motivating factors. Across all ages, only a small minority report lacking understanding of the importance of PA. 

# Statistical Analysis 

Multigroup SEM was used to to compare mean levels of PA motives between adults and youths, followed by LCA to identify distinct motivational profiles within each group. Class membership was then regressed on age group using multinomial logistic regression, allowing for a direct comparison of age-related patterns in profile membership between adults and youths. Two motive items (challenge and importance) were excluded from multigroup SEM due to non-equivalent wording across the surveys but were retained for LCA. Data used for the two analyses were similar in terms of gender and ethnicity distributions. 

## Multigroup Structural Equation Modeling

Differences in the slopes of motive items were examined while controlling for demographic factors (see Appendix E for model specification). This approach directly addresses H1(a), H1(b), and H1(c) by testing whether the strength of association between each motive item and weekly PA varies between youths and adults. 

```{r summary.bi, fig.align='center',tab.cap='Proportion of "strongly agree" responses for adults (left) and youths (right).'}
opts <- options(knitr.kable.NA = '')
kable(list(adult.summary.bi, child.summary.bi),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

To account for differences in Likert scales between adults and youths, all motive items were dichotomized into strongly agree and not strongly agree. As shown in *Table 1*, the distributions of positive responses exhibit a left skew in both adults and youths, suggesting that participants were less likely to choose the highest agreement category across items. A cap of 1,680 minutes per week was applied to reported PA to minimize the influence of extreme values or potential data entry errors. `r prettyNum(nrow(adult.bi), big.mark=",")` adult and `r prettyNum(nrow(child.bi), big.mark=",")` youth respondents were included. 
A freely estimated model was first fitted, followed by a series of constrained models in which individual or all motive pathways were fixed to equality across groups. This stepwise approach allowed assessment of whether the strength of specific motives differed significantly between youths and adults. An ANOVA was then conducted to test the significance of differences between models. Ethnicity, gender, and age were controlled for in all models.

Prior to fitting the multigroup SEM models, assumptions were assessed. Moderate correlations were observed among the motive items in both youth and adult samples, with most values ranging between 0.2 and 0.5. The highest correlations were between enjoyment and ability in youths (r = 0.656) and between enjoyment and perceived importance in adults (r = 0.674), slightly exceeding the conventional threshold of r = 0.6 for concern about multicollinearity. Given the conceptual similarity of these motive items, a level of correlation is expected and largely unavoidable. Standardized residual correlations were also examined to evaluate local fit for both youth and adult models. Since all values were well below 0.7, all items can be retained (Che Lah & Tasir, 2018). The largest residuals occurred among closely related motive items (enjoyment, fitness, opportunity, and relaxation), exceeding 0.4. This misfit likely reflects conceptual overlap and the use of dichotomized indicators. Nevertheless, residuals involving demographics and PA minutes were consistently low. These localized discrepancies do not affect the primary purpose, which is to test whether the strengths of associations, rather than the predictive power of motives on PA, differ across groups.

## Latent Profile Analysis

LCA was conducted separately for youths and adults to identify distinct motivational patterns, using the Likert-scale motive items as indicators. Given the limited number of motive items and the use of a single item to measure each construct, profiles are interpreted descriptively based on relative endorsement patterns rather than being formally mapped onto theoretical constructs. Ethnicity and gender were included as covariates predicting class membership. Multiple sets of random starting values were used to ensure solution stability and convergence. The final sample included `r prettyNum(nrow(adult.lik), big.mark=",")` adult and `r prettyNum(nrow(child.lik), big.mark=",")` youth respondents.

The optimal number of classes was determined based on AIC, BIC, relative entropy, the Lo-Mendell-Rubin likelihood ratio test (LMR), the bootstrap likelihood ratio test (BLRT), class proportions, and substantive interpretability. Multinomial logistic regression with age as a predictor was then applied to examine the relationship between age and profile membership, with odds ratios and 95% confidence intervals calculated by exponentiating the estimated coefficients. This approach directly tests whether the probability of belonging to lower motivation profiles increases with age, addressing H2(a) for adults and H2(b) for youths.

Conditional independence among PA-related items was assessed using standardized bivariate chi-square statistics. Extremely sparse cells (<1% of observations) inflated these values, making them unreliable indicators of local dependence. Given that the key motive items represent conceptually distinct constructs, minor violations are less concerning than for highly overlapping indicators. Model evaluation relied on conventional LCA fit indices, including BIC and entropy, alongside substantive interpretability. These criteria indicated that the identified profiles were robust and meaningful despite the limitations of the conditional independence tests. 

## Software and Packages

All analyses were conducted in R (version 4.4.3). Multigroup SEM was performed using the lavaan package (Rosseel, 2012; Rosseel *et al.*, 2025). LCA and related statistics were conducted using packages poLCA (Linzer & Lewis, 2011) and poLCAExtra (Caron, 2024).

# Results & Discussion
  
## Multigroup SEM

Significant differences between adults and youths were observed in the strength of associations between motive items and weekly PA. The ANOVA results showed that each constrained model fit significantly worse than the freely estimated model, indicating that allowing all associations to vary by age group provided the best representation of the data. Coefficients in *Table 2* represent the estimated change in weekly PA minutes associated with strong agreement for each binary motive. 

```{r slopes.diff, fig.align='center',tab.cap="Estimated slopes of each motivational factor in the SEM, with differences calculated as youth minus adult values. All differences were statistically significant (p < 0.05)."}
opts <- options(knitr.kable.NA = '')
kable(list(slopes.diff),align='c',booktabs = T) %>% 
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

Though modest in magnitude, enjoyment and guilt were more strongly associated with PA in youths, supporting H1(a). By contrast, the social motive was stronger in adults instead. The observed discrepancy may be attributed to several methodological and conceptual factors. Variations in survey question wording could influence how respondents interpret and respond to items assessing social motives. For instance, Geller *et al.* (2018) defines social motives as the extrinsic drive to socially interact and meet new people during PA, emphasizing the external nature of this motivation. In contrast, Bragg *et al.* (2018) categorizes social motive as social influence, encompassing the impact of friends, parents, family members, coaches, and health professionals on an individual's PA choices. These differing definitions may lead to variations in how social motives are perceived and reported across studies. In addition, the conflation of various forms of PA in survey instruments can obscure the specific motives associated with different activities. For example, Vuckovic & Duric (2024) found that social motives were less influential in fitness activities but more significant in sport participation. However, in surveys that combines all types of PA without distinguishing between them, such as the one used for this dissertation, the unique impact of social motives on each activity type may be masked.

Fitness was more strongly associated with PA in adults than in youths, as expected, whereas relaxation was more influential in youths, contrary to H1(b). Several developmental and contextual factors may provide insight. Not only was achieving relaxation one of the most commonly reported motives for participation of PA in adolescents aged 16 and older (Ashford *et al.*, 1993), it's also a key determinant of fun in youth sports that promotes sustained engagement (Visek *et al.*, 2015). In addition, Teuber *et al.* (2024) reported that engagement in PA not only reduced stress but also enhanced perceived academic performance in university students. While this sample is older than the youths in the present study, the academic pressure context is analogous. On the other hand, elevated momentary stress was found to predict lower subsequent PA (Do *et al.*, 2021). Since the present study does not assess causality, there is the possibility that relaxation drive higher PA rather than the reverse. More broadly, personal experiences appear to foster understanding of the benefits of PA, which may in turn drive PA engagement (Kostamo *et al.*, 2019). Although relaxation serves as a more prominent motive for females (Kondric *et al.*, 2013), the fairly balanced gender distribution in our youth sample suggests this factor does not meaningfully influence the overall pattern observed.

Contrary to H1(c), environmental opportunity was more strongly associated with PA among adults, with the biggest difference among all motives at 63 minutes. This discrepancy may be due to youths having more consistent access to facilities such as school gyms and playgrounds, while adults are limited by the rising opportunity costs associated with PA participation. This is particularly true among highly educated adults who are aware of the health benefits, but often hold sedentary (Brown & Roberts, 2011). Furthermore, women often face fewer chances to engage in PA than men, particularly in cultural contexts where they are primarily responsible for household caretaking, which restricts access to outdoor or leisure-based activity (Jalaluddin *et al.*, 2024). 

These findings demonstrate clear shifts in motivational profiles. Youths appear more responsive to internalized motives, whereas external factors play a larger role in adults. Notably, the difference associated with opportunity is far larger than that of any other motive, suggesting that adult engagement in PA is strongly contingent on factors such as access, time, and environmental support. Within the SDT framework, this would suggest that as youths transition to adulthood, intrinsic, identified, and introjected regulation tend to weaken. The social motive may be an exception, serving to satisfy intrinsic needs that are less fulfilled in adulthood. Alternatively, it may be interpreted that the determinants of intention appear to shift with age. In the context of the TPB, the results imply personal valuation and perceived social expectations have a stronger influence on youths and decreases with age. However, despite these patterns, the differences for most motives are weak to moderate in magnitude, suggesting that the overall motivational structure may be largely similar.

## LCA

### Adult Classes

```{r elbow.llik.ad, out.width='50%',fig.show='hold',fig.align='center', fig.cap='Elbow curve of AIC and BIC values across different numbers of latent classes in adults.'}

gg.elbow.ad

```

Latent class models ranging from 3 to 6 classes all demonstrated low AIC and BIC values (Figure 5), significant LMR and BLRT results, and reasonable average posterior entropy per class. Starting form the 4-class model, there appears to be diminishing return, suggesting that adding more than 4 classes did not provide a substantial improvement in model fit. While both the 3-class and 4-class models have adequate relative entropy values at  `r LCAE.ad$output[2,6]` and `r LCAE.ad$output[3,6]`, respectively, the 4-class solution contains two groups with nearly identical response patterns, which undermines substantive interpretation. The bootstrap likelihood ratio test was used to assess the significance of performance differences. All models except the 7-class were found to be significant, with a preference for the 6-class solution. Nevertheless, the 3-class model is selected for its balance between simplicity and the ability to distinguish meaningful motivational profiles. Fit indices for all models can be found in Appendix D.


```{r distr.ad, fig.width="70%", fig.show='hold',fig.align='center', fig.cap='Frequency of adult responses for each motive by class membership.'}

plot(LCAE.ad, nclass = 2, main = "Response Distributions by Class, Adults")

```

Although distinct profiles were identified, they largely corresponded to three general levels of motivation, limiting the differentiation of more nuanced patterns. *Figure 6* shows the distribution of responses across each latent class.

**Class 1** represents adults who consistently report strong agreement across a broad range of motivations. Members show high probabilities of strongly agreeing with enjoyment, fitness, ability, opportunity, importance, and relaxation, and moderately agreeing with social, challenge, and guilt motives. This group is labeled the High Motivation class.

**Class 2** is characterized by very low probabilities of strongly agreeing with opportunity and ability, and by a higher prevalence of neutral or negative responses across most other items. Members tend to express the least favorable attitudes toward PA. This group is labelled the Low Motivation class. It represents a relatively small proportion of the sample (~6%), which corresponds to the patterns observed in the raw data, where very few respondents selected negative responses in general.

**Class 3** exhibits a generally positive attitude toward various motives across the board. These individuals are less likely to show extreme positive endorsement than class 1, and are labeled the Moderate Motivation class.

```{r mins.ad, out.width='50%', fig.show='hold',fig.align='center', fig.cap='Weighted median and interquartile range of minutes of PA per week for adults by class.'}

gg.mins.ad

```
 
There exists substantial differences in the median of weekly PA minutes, as seen in *Figure 7*. The High Motivation class exhibits the highest levels (445), followed by the Moderate Motivation class (182.5), and finally the Low Motivation class (45). Notably, the interquartile range is positively associated with activity level, highlighting considerable variability, particularly among the most highly active class. 

```{r byage.ad, out.width='50%',fig.show='hold',fig.align='center', fig.cap='Proportion of adults in each class across age groups.'}

gg.byage.ad

```

The distribution of age groups highlights notable patterns in motivational profiles (*Figure 8*). The High Motivation class is consistently the second most populous across age groups, gradually declining with increasing age. The Low Motivation class remains the smallest in all age groups, with its relative size largely stable across ages. In contrast, the Moderate Motivation class contains the largest number of adults in every age group, with membership increasing steadily as age rises, suggesting that some individuals may shift from high to moderate motivation over time.

```{r or.ad, fig.align='center',tab.cap="Odds ratios for adult class membership across age groups, with the youngest group (16-34 years) and High Motivation class as the reference."}
opts <- options(knitr.kable.NA = '')
kable(list(or.ad),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

Multinomial logistic regression was applied to directly examine the relatioship between age groups and class membership, with the High Motivation class and youngest group (16-34) as reference (*Table 3*). The youngest adults are more than 6.6 times as likely to belong to the High Motivation class compared to the Low Motivation class. Although their odds of falling into the Moderate Motivation group are 15% higher, this may be due to the larger size of that group. For the Low Motivation category, no statistically significant differences are found between the 35-44 and 55-64 age groups and the reference category. Other age groups show similar odds, except for those aged 75 and above, who are 1.5 times more likely to be in the Low Motivation class. Meanwhile, the likelihood of being in the Moderate Motivation group rises with age, with adults aged 65-74 being 1.45 times and those aged 75+ being 1.93 times more likely than the youngest adults. While class membership remains relatively stable among adults under 65, those older than 65 are more likely to be classified within low or moderate motivation groups, consistent with H2(a).

The marked reduction in motivation and PA observed among older adults may stem from the interaction of perceived physical limitations and changing psychosocial influences. Advancing age is frequently accompanied by beliefs of inevitable physical decline, which can erode perceptions of competence and autonomy, weakening intrinsic motivation (Palombi *et al.*, 2025). Such beliefs may also act as self-fulfilling constraints, discouraging engagement even when actual physical capacity remains sufficient. At the same time, negative physiological experiences such as pain or fatigue, combined with reduced PBC over PA outcomes, further diminish favourable attitudes and behavioural intentions. Traditional social motivators, such as conformity to group norms or identification with active peers, become less influential (Meredith *et al*., 2023; Tsai *et al.*, 2022), suggesting that subjective norm and external pressures play a reduced role in driving this age-related decline in motivation. 


### Youth Classes
 
```{r elbow.llik.ch, out.width='50%',fig.show='hold',fig.align='center', fig.cap='Elbow curve of AIC and BIC values across different numbers of latent classes in youths.'}

gg.elbow.ch

```

Increasing the number of classes beyond 4 led to a decrease, as evidenced by a sharp increase in AIC and BIC (*Figure 9*). The penalty for model complexity likely outweighed any improvements in fit. Consequently, the 3-class and 4-class solutions were the only viable candidates. The LMR and BLRT results further support this, indicating significance only for these two models, with a preference for the 4-class solution. Both models show acceptable relative entropy values at `r LCAE.ch$output[2,6]` and `r LCAE.ch$output[3,6]`, but the higher relative entropy for the 3-class model suggests it offers a more nuanced classification of individuals. Similar to the adults, the 4-class model also resulted in two very similar classes, suggesting that the additional class did not add meaningful differentiation. Given these considerations, the 3-class model was selected


```{r distr.ch, out.width='70%',fig.show='as.is',fig.align='center', fig.cap='Frequency of youth responses for each motive by class membership.'}

plot(LCAE.ch, nclass = 2, main = "Response Distributions by Class, Youths")

```

The LCA produced profiles similar to those observed in adults, largely reflecting three general motivation levels rather than more nuanced distinctions. *Figure 10* illustrates the response patterns for class.

**Class 1** represents youths who consistently report positive attitudes toward a wide range of motives. Members show high probabilities of strongly agreeing with motives related to enjoyment, fitness, opportunity, importance, and ability. They also moderately endorse social, challenge, and relaxation motives, and express relatively high agreement with guilt. This class is labeled the High Motivation class.

**Class 2** is characterized by very low probabilities of strongly agreeing with any motives, and by a higher prevalence of neutral or negative responses across most other items. Members tend to express the least favorable attitudes toward PA. This class is labeled the Low Motivation class, and represents a relatively small proportion of the sample. This is consistent with the raw data, where few respondents selected negative responses.


**Class 3** exhibits a moderate attitude toward most motives, with moderate positive endorsements of enjoyment, fitness, opportunity, ability, and importance. Disagreement and strong agreement are both uncommon. This group falls in the middle, representing generally positive but not strongly emphatic motivation toward PA. This class is labeled the Moderate Motivation class.

```{r mins.ch, out.width='50%', fig.show='as.is',fig.align='center', fig.cap='Weighted median and interquartile range of minutes of PA per week for youths by class.'}

gg.mins.ch

```


Similar to adults, significant differences in the median weekly PA minutes are observed across the classes, along with substantial variability within each class (*Figure 11*). The High Motivation class boasts the highest number (405) and highest interquartile range, followed by the Moderate Motivation class (235), and then the Low Motivation class(125). 

```{r byage.ch, out.width='50%',fig.show='hold',fig.align='center', fig.cap='Proportion of youths in each class by age.'}

gg.byage.ch

```

Figure 12 highlights notable differences in the distribution of age. The High and Moderate Motivation classes are roughly similar in size, with the latter decreasing with age, except for the sudden increase at age 16. The Low Motivation class is smaller than the other two but still larger than in the adult sample. It steadily gains size with age, with a slight dip at age 16. These patterns suggest that while motivation generally declines during youth, a significant portion of youths maintain moderate to strong motivation throughout. Additionally, some youths may experience a boost in motivation during late adolescence.


```{r or.ch, fig.align='center',tab.cap="Odds ratios for youth class membership across ages, with 11-year-olds and High Motivation class as the reference."}
opts <- options(knitr.kable.NA = '')
kable(list(or.ch),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```


The multinomial logistic regression revealed statistically significant differences across all age groups and classes, with the reference group being 11-year-olds in the High Motivation class (*Table 4*). At age 11, youths are approximately 5 times less likely to belong to the Low Motivation class compared to the reference group, while their odds of belonging to the Moderate Motivation class are about the same. For the Low Motivation class, the odds of membership increase steadily with age, peaking at age 15, before slightly decreasing at age 16, from twice as likely to 1.36 times as likely. This supports the pattern observed in *Figure 12*, suggesting that these individuals may have shifted to the High Motivation class. In contrast, the odds for the Moderate Motivation class remain relatively stable across ages, with a slight decrease at age 11 and 16. These findings highlight a shift toward lower motivation as youths age, with the slight exception of 16-year-olds, generally supporting H2(b).

There is a steady and substantial increase (20-30% per year) in the likelihood of belonging to the low motivation class, a finding consistent with prior research (Shao & Zhou, 2023; Mendonça *et al.*, 2014). This trend aligns with prior observations that younger youths are more motivated by mastery goals, which emphasise learning and competence and are positively associated with PA levels, compared to older adolescents (Sahin *et al.*, 2016). Conversely, performance-avoidance goals, which increase with age, are negatively associated with enjoyment (Barkoukis *et al.*, 2024). As the motivational focus shifts from mastery to avoidance-oriented goals, enjoyment and intrinsic motivation diminish as a result (Guan *et al.*, 2023), which may underlie the observed increase in low motivation profiles with age. Additionally, parental modelling and logistical support decline with age, despite their crucial role in fostering PA habits   (Davison & Jago, 2009). This may also account for the slowing in the overall decline of motivation observed around age 16. At this stage, youths gain greater autonomy and become less dependent on adult support, which can foster a renewed increase in intrinsic motivation. Concurrent legal and social transitions, such as eligibility for independent travel or part-time work, further expand opportunities for participation.

### Cross-Group Comparison

While profiles for both adults and youths reflect general motivation levels rather than distinct configurations of motives, they exhibit broadly consistent patterns. In both groups, the youngest individuals are highly unlikely to belong to lower motivation profiles. Among adults, profile membership remains relatively stable through middle age, with a pronounced shift toward lower motivation emerging only in later life. This plateau may reflect the relative stability of daily routines, responsibilities, and health prior to older adulthood. In contrast, among youths, membership in low and moderate motivation profiles increases steadily with age, with low motivation rising more sharply.

These patterns suggest that age is positively correlated with motivational profiles similarly across developmental stages, but the mechanisms and timing differ. In youths, the gradual increase in lower motivation may result from accumulating social pressures, heightened self consciousness, and emerging autonomy that introduces variability in engagement. Among older adults, the abrupt rise in low motivation likely reflects structural and physiological transitions, including retirement, health decline, or reduced access to organized PA. While the slope and onset differ, both groups exhibit an age-related shift toward low motivation profiles, supporting the broader hypothesis that age shapes motivational profile membership similarly across developmental stages. 

# Conclusion

This study examined how perceived PA motives and PA vary across developmental stages and how age shapes dominant motives within youth and adult groups. Specifically, it addressed the research questions: (1) How does the relationship between perceived motives and PA differ between youths and adults? and (2) What is the relationship between age and motivational profiles of PA in youths and adults?

The findings indicate that motives and PA are related differently across developmental stages, although the patterns do not fully align with initial predictions. Social motives were unexpectedly stronger among adults than youths, and relaxation motives were stronger in youths, indicating that H1(a) and H1(b) are only partially supported. Contrary to H1(c), perceived opportunity was more strongly associated with PA in adults than in youths, suggesting that environmental and structural factors may constrain adult activity more than previously assumed. 

H2(a) and H2(b) received more consistent support. Distinct motivational profiles were identified within both youths and adults, and age was associated with systematic shifts in profile membership in both groups. Among adults, motivation gradually declined with age, with the oldest adults exhibiting the steepest drop toward Low Motivation profiles, consistent with H2(a). Among youths, High Motivation class membership also declined gradually, but the oldest adolescents showed a sudden increase, providing only partial support for H2(b). 

Overall, the results show that the associations between perceived motives and PA differ between youths and adults, and that age and motivational profiles also vary across groups in similar ways. However, these patterns are complex and, in some cases, deviate from theoretical predictions.

## Policy Implications

For adults, structural and social opportunities are strongly associated with PA. Workplace initiatives, flexible scheduling, and accessible community facilities can reduce the opportunity cost of PA for those balancing work, family, and other responsibilities. As intrinsic motives such as enjoyment may decline with age, social motives may help maintain PA by providing extrinsic incentives. Policies should also address differential access, particularly for women with domestic responsibilities, by providing inclusive, low barrier options. In older adulthood, declines in intrinsic motivation and physical capacity necessitate targeted environmental and social support. 

For youths, providing choice in activity type, intensity, and scheduling can help sustain engagement, particularly as enjoyment and personally meaningful goals are primary motivators. Fun, low pressure activities remain essential, given that relaxation and stress relief are consistently valued motives. Adult support should focus on logistical assistance and modelling rather than mastery or performance pressure, as avoidance tendencies naturally increase with age. Around age 16, when adolescents gain greater control over their schedules and become less reliant on adult support, policies should facilitate access to resources, guidance, and safe environments for self-directed activity, helping translate increased autonomy into sustained intrinsic participation.

## Limitations

Several limitations may affect the interpretation of these results. All data were self-reported, making them susceptible to recall errors, comprehension difficulties, and social desirability bias, particularly among students responding under teacher supervision. The adult sample, collected primarily online, may underrepresent individuals with limited digital literacy or internet access, such as older adults or low income households. Allowing up to two respondents per household could have increased intra-household similarity, and minor differences between online and paper formats may have influenced reported PA or demographic details.

A major limitation is that the data were not collected using an established, validated questionnaire (e.g., BREQ-2), so corresponding theoretical constructs were not formally operationalized. Each variable relied on a single-item response, potentially limiting reliability and validity. In addition, broader SDT dimensions were not captured, meaning latent class profiles cannot be conclusively labelled as "self-determined" or "autonomous" While they reflect relative endorsement patterns, any interpretation in terms of autonomy or controlled motivation remains speculative and descriptive.

To ensure comparability across groups, only motive items with identical or near-identical wording were retained, which may have omitted relevant constructs. PA measures were highly skewed, and analyses focused on moderate-to-vigorous activity, potentially underestimating meaningful engagement among older adults who benefit from light activity.

Finally, demographic factors such as socioeconomic status, BMI, education, and relationship status were not considered, and the cross-sectional design prevents causal inference about developmental changes in motivation or behaviour.

\newpage

# References

Ajzen, I. (1991). 'The theory of planned behavior', *Organizational Behavior and Human Decision Processes*, 50(2), 179-211. [https://doi.org/10.1016/0749-5978(91)90020-T](https://doi.org/10.1016/0749-5978(91)90020-T).

Ajzen, I. (2020). 'The theory of planned behavior: Frequently asked questions', *Human Behavior and Emerging Technologies*, 2(4), 314-324. [https://doi.org/10.1002/hbe2.195](https://doi.org/10.1002/hbe2.195).

Anderson, E., & Shivakumar, G. (2013). 'Effects of exercise and physical activity on anxiety', *Frontiers in Psychiatry*, 4, 27. [https://doi.org/10.3389/fpsyt.2013.00027](https://doi.org/10.3389/fpsyt.2013.00027).

Ashford, B., Biddle, S., & Goudas, M. (1993). 'Participation in community sports centres: motives and predictors of enjoyment', *Journal of Sports Sciences*, 11(3), 249-256. [https://doi.org/10.1080/02640419308729992](https://doi.org/10.1080/02640419308729992).

Barkoukis, V., Gråstén, A., Huhtiniemi, M., & Jaakkola, T. (2024). 'Developmental relations of achievement goals and affect in physical education', *Psychology of Sport and Exercise*, 75, 102700. [https://doi.org/10.1016/j.psychsport.2024.102700](https://doi.org/10.1016/j.psychsport.2024.102700).

Biddle, S. J. H., Atkin, A. J., Cavill, N., & Foster, C. (2011). 'Correlates of physical activity in youth: a review of quantitative systematic reviews', *International Review of Sport and Exercise Psychology*, 4(1), 25-49. [http://dx.doi.org/10.1080/1750984X.2010.548528](http://dx.doi.org/10.1080/1750984X.2010.548528).

Booth, F. W., Roberts, C. K., & Laye, M. J. (2012). 'Lack of exercise is a major cause of chronic diseases', *Comprehensive Physiology*, 2(2), 1143-1211. [https://doi.org/10.1002/cphy.c110025](https://doi.org/10.1002/cphy.c110025).

Bragg, M. A., Tucker, C. M., Kaye, L. B., & Desmond, F. (2009). 'Motivators of and barriers to engaging in physical activity: Perspectives of low-income culturally diverse adolescents and adults', *American Journal of Health Education*, 40(3), 146-154. [https://doi.org/10.1080/19325037.2009.10599089](https://doi.org/10.1080/19325037.2009.10599089).

Brown, H., & Roberts, J. (2011). 'Exercising choice: the economic determinants of physical activity behaviour of an employed population', *Social Science & Medicine*, 73(3), 383-390. [https://doi.org/10.1016/j.socscimed.2011.06.001](https://doi.org/10.1016/j.socscimed.2011.06.001).

Brooks, J. M., Iwanaga, K., Chiu, C. Y., Cotton, B. P., Deiches, J., Morrison, B., Moser, E., & Chan, F. (2017). 'Relationships between self-determination theory and theory of planned behavior applied to physical activity and exercise behavior in chronic pain', *Psychology, Health & Medicine*, 22(7), 814-822. [https://doi.org/10.1080/13548506.2017.1282161](https://doi.org/10.1080/13548506.2017.1282161).

Brunet, J., & Sabiston, C. M. (2011). 'Exploring motivation for physical activity across the adult lifespan', *Psychology of Sport and Exercise*, 12(2), 99-105. [https://doi.org/10.1016/j.psychsport.2010.09.006](https://doi.org/10.1016/j.psychsport.2010.09.006).

Caron, P.-O. (2024). 'poLCAExtra: New and convenient functions for the package 'poLCA''. [https://github.com/quantmeth/poLCAExtra](https://github.com/quantmeth/poLCAExtra).

Che Lah, N. H., & Tasir, Z. (2018). Measuring reliability and validity of questionnaire on online social presence: A Rasch model analysis. *Journal of Computational and Theoretical Nanoscience, 24*(11), 7900-7903. [https://doi.org/10.1166/asl.2018.12452](https://doi.org/10.1166/asl.2018.12452).

Choi, J., Lee, M., Lee, J. K., Kang, D., Choi, J-Y. (2017). 'Correlates associated with participation in physical activity among adults: a systematic review of reviews and update', *BMC Public Health*, 17, 356. [https://doi.org/10.1186/s12889-017-4255-2](https://doi.org/10.1186/s12889-017-4255-2).

Daley, A. J., & Duda, J. L. (2006). 'Self-determination, stage of readiness to change for exercise, and frequency of physical activity in young people', *European Journal of Sport Science*, 6(4), 231-243. [https://doi.org/10.1080/17461390601012637](https://doi.org/10.1080/17461390601012637).

Davison, K. K., & Jago, R. (2009). 'Change in parent and peer support across ages 9 to 15 yr and adolescent girls' physical activity', *Medicine and Science in Sports and Exercise*, 41(9), 1816-1825. [https://doi.org/10.1249/MSS.0b013e3181a278e2](https://doi.org/10.1249/MSS.0b013e3181a278e2).

de Maio Nascimento, M., Gouveia, É. R., Gouveia, B. R., Marques, A., França, C., Campos, P., Martins, F., García-Mayor, J. & Ihle, A. (2023). 'Differential patterns in motivations for practising sport and their effects on physical activity engagement across the lifespan', *Healthcare (Basel, Switzerland)*, 11(2), 274. [https://doi.org/10.3390/healthcare11020274](https://doi.org/10.3390/healthcare11020274).

Deci, E. L., & Ryan, R. M. (2008). 'Self-Determination Theory: A macrotheory of human motivation, development, and health', *Canadian Psychology/Psychologie Canadienne*, 49, 182-185. [https://doi.org/10.1037/a0012801](https://doi.org/10.1037/a0012801).

Dishman, R. K., McIver, K. L., Dowda, M., & Pate, R. R. (2018). 'Declining physical activity and motivation from middle school to high school', *Medicine and Science in Sports and Exercise*, 50(6), 1206-1215. [https://doi.org/10.1249/MSS.0000000000001542](https://doi.org/10.1249/MSS.0000000000001542).

Do, B., Mason, T. B., Yi, L., Yang, C. H., & Dunton, G. F. (2021). 'Momentary associations between stress and physical activity among children using ecological momentary assessment', *Psychology of Sport and Exercise*, 55, 101935. [https://doi.org/10.1016/j.psychsport.2021.101935](https://doi.org/10.1016/j.psychsport.2021.101935).

Duncan, L. R., Hall, C. R., Wilson, P. M., & Wilson, M. D. (2010). 'Exercise motivation: a cross-sectional analysis examining its relationships with frequency, intensity, and duration of exercise', *International Journal of Behavioral Nutrition and Physical Activity*, 7(7). [https://doi.org/10.1186/1479-5868-7-7](https://doi.org/10.1186/1479-5868-7-7).

Geller, K., Renneke, K., Custer, S., & Tigue, G. (2018). 'Intrinsic and extrinsic motives support adults' regular physical activity maintenance', *Sports Medicine International Open*, 2(3), E62-E66. [https://doi.org/10.1055/a-0620-9137](https://doi.org/10.1055/a-0620-9137).

Godin, G., & Kok, G. (1996). 'The theory of planned behavior: a review of its applications to health-related behaviors', *American Journal of Health Promotion*, 11(2), 87-98. [https://doi.org/10.4278/0890-1171-11.2.87](https://doi.org/10.4278/0890-1171-11.2.87).

Guan, J., Xiang, P., Land, W., & Hamilton, X. D. (2023). 'Age and gender differences in achievement goal orientations in relation to physical activity', *Perceptual and Motor Skills*, 130(1), 80-93. [https://doi.org/10.1177/00315125221139000](https://doi.org/10.1177/00315125221139000).

Hallal, P. C., Andersen, L. B., Bull, F. C., Guthold, R., Haskell, W., & Ekelund, U. (2012). 'Global physical activity levels: surveillance progress, pitfalls, and prospects', *Lancet*, 380(9838), 247-257. [https://doi.org/10.1016/S0140-6736(12)60646-1](https://doi.org/10.1016/S0140-6736(12)60646-1).

Hagger, M. S., Chatzisarantis, N. L. D., & Biddle, S. J. H. (2002). 'A meta-analytic review of the theories of reasoned action and planned behaviour in physical activity: Predictive validity and the contribution to additional variables', *Journal of Sport and Exercise Psychology*, 24, 3-32. [https://doi.org/10.1123/jsep.24.1.3](https://doi.org/10.1123/jsep.24.1.3).

Heredia-León, D. A., Valero-Valenzuela, A., Gómez-Mármol, A., & Manzano-Sánchez, D. (2023). 'Motivational profiles in physical education: Differences at the psychosocial, gender, age and extracurricular sports practice levels', *Children*, 10(1), 112. [https://doi.org/10.3390/children10010112](https://doi.org/10.3390/children10010112).

Jalaluddin, N., Khalique, N., & Ahmad, A. (2024). 'Physical inactivity and its sociodemographic determinants among adults: A community-based study', *Indian Journal of Community Medicine*, 49(6), 849-854. [https://doi.org/10.4103/ijcm.ijcm_130_23](https://doi.org/10.4103/ijcm.ijcm_130_23).

Kim, J., Eys, M., Robertson-Wilson, J., Dunn, E., & Rellinger, K. (2019). 'Subjective norms matter for physical activity intentions more than previously thought: Reconsidering measurement and analytical approaches', *Psychology of Sport and Exercise*, 43, 359-367. [https://doi.org/10.1016/j.psychsport.2019.04.013](https://doi.org/10.1016/j.psychsport.2019.04.013).

Kondric, M., Sindik, J., Furjan-Mandić, G., & Schiefler, B. (2013). 'Participation motivation and students' physical activity among sport students in three countries', *Journal of Sports Science & Medicine*, 12(1), 10-18. [https://pmc.ncbi.nlm.nih.gov/articles/PMC3761756/](https://pmc.ncbi.nlm.nih.gov/articles/PMC3761756/). Accessed: 20 October 2025.

Kostamo, K., Vesala, K. M., & Hankonen, N. (2019). 'What triggers changes in adolescents' physical activity? Analysis of critical incidents during childhood and youth in student writings', *Psychology of Sport and Exercise*, 45, 101564. [https://doi.org/10.1016/j.psychsport.2019.101564](https://doi.org/10.1016/j.psychsport.2019.101564).

Lee, I. M., Shiroma, E. J., Lobelo, F., Puska, P., Blair, S. N., & Katzmarzyk, P. T. (2012). 'Effect of physical inactivity on major non-communicable diseases worldwide: an analysis of burden of disease and life expectancy', *Lancet*, 380(9838), 219-229. [https://doi.org/10.1016/S0140-6736(12)61031-9](https://doi.org/10.1016/S0140-6736(12)61031-9).

Linzer, D. A., & Lewis, J. B. (2011). 'poLCA: An R package for polytomous variable latent class analysis', *Journal of Statistical Software*, 42(10), 1-29. [https://www.jstatsoft.org/v42/i10/](https://www.jstatsoft.org/v42/i10/).

Mendonça, G., Cheng, L. A., Mélo, E. N., & de Farias Júnior, J. C. (2014). 'Physical activity and social support in adolescents: a systematic review', *Health Education Research*, 29(5), 822-839. [https://doi.org/10.1093/her/cyu017](https://doi.org/10.1093/her/cyu017).

Meredith, S. J., Cox, N. J., Ibrahim, K., Higson, J., McNiff, J., Mitchell, S., Rutherford, M., Wijayendran, A., Shenkin, S. D., Kilgour, A. H. M., & Lim, S. E. R. (2023). 'Factors that influence older adults' participation in physical activity: a systematic review of qualitative studies', *Age and Ageing*, 52(8), afad145. [https://doi.org/10.1093/ageing/afad145](https://doi.org/10.1093/ageing/afad145).

Mitchell, J. (2019). 'Physical inactivity in childhood from preschool to adolescence', *ACSM's Health & Fitness Journal*, 23(5), 21-25. [https://doi.org/10.1249/fit.0000000000000507](https://doi.org/10.1249/fit.0000000000000507).

Moreno-Murcia, J. A., Cervelló Gimeno, E., Hernández, E. H., Pedreño, N. B., & Rodríguez Marín, J. J. (2013). 'Motivational profiles in physical education and their relation to the theory of planned behavior', *Journal of Sports Science & Medicine*, 12(3), 551-558. [https://pmc.ncbi.nlm.nih.gov/articles/PMC3772601](https://pmc.ncbi.nlm.nih.gov/articles/PMC3772601). Accessed: 12 October 2025.

Neipp, M., Quiles, M., Leon, E., & Rodríguez-Marín, J. (2013). 'Theory of Planned Behavior and physical exercise: Differences between people who do regular physical exercise and those who do not', *Wulfenia*, 20, 324-335. [https://www.researchgate.net/publication/262680419_Theory_of_Planned_Behavior_and_physical_exercise_Differences_between_people_who_do_regular_physical_exercise_and_those_who_do_not](https://www.researchgate.net/publication/262680419_Theory_of_Planned_Behavior_and_physical_exercise_Differences_between_people_who_do_regular_physical_exercise_and_those_who_do_not). Accessed: 15 August 2025.

Ng, J. Y., Ntoumanis, N., Thøgersen-Ntoumani, C., Deci, E. L., Ryan, R. M., Duda, J. L., & Williams, G. C. (2012). 'Self-determination theory applied to health contexts: A meta-analysis', *Perspectives on Psychological Science*, 7(4), 325-340. [https://doi.org/10.1177/1745691612447309](https://doi.org/10.1177/1745691612447309).

Nuss, K., Sui, W., Rhodes, R., & Liu, S. (2023). 'Motivational profiles and associations with physical activity before, during, and after the COVID-19 pandemic: Retrospective study', *JMIR Formative Research*, 7, e43411. [https://doi.org/10.2196/43411](https://doi.org/10.2196/43411).

Ostendorf, D. M., Schmiege, S. J., Conroy, D. E., Phelan, S., Bryan., A. D., Catenacci, V. A. (2021). 'Motivational profiles and change in physical activity during a weight loss intervention: a secondary data analysis', *International Journal of Behavioral Nutrition and Physical Activity*, 18, 158. [https://doi.org/10.1186/s12966-021-01225-5](https://doi.org/10.1186/s12966-021-01225-5).

Oyibo, K., Adaji, I., & Vassileva, J. (2018). 'Social cognitive determinants of exercise behavior in the context of behavior modeling: a mixed method approach', *Digital Health*, 4, 2055207618811555. [https://doi.org/10.1177/2055207618811555](https://doi.org/10.1177/2055207618811555).

Palombi, T., Chirico, A., Cazzolli, B., Zacchilli, M., Alessandri, G., Filosa, L., Borghi, A., Fini, C., Antoniucci, C., Pistella, J., Alivernini, F., Baiocco, R., & Lucidi, F. (2025). 'Motivation, psychological needs and physical activity in older adults: a qualitative review', *Age and Ageing*, 54(7), afaf180. [https://doi.org/10.1093/ageing/afaf180](https://doi.org/10.1093/ageing/afaf180).

Reeve, J. (2012). A self-determination theory perspective on student engagement. In S. L. Christenson, A. L. Reschly, & C. Wylie (Eds.), *Handbook of research on student engagement (pp. 149-172)*. Springer. [https://doi.org/10.1007/978-1-4614-2018-7_7](https://doi.org/10.1007/978-1-4614-2018-7_7).

Rosseel, Y. (2012). 'lavaan: An R package for structural equation modeling', *Journal of Statistical Software*, 48(2), 1-36. [https://doi.org/10.18637/jss.v048.i02](https://doi.org/10.18637/jss.v048.i02).

Rosseel, Y., Jorgensen, T., & De Wilde, L. (2025). 'lavaan: Latent variable analysis', R package version 0.6-20. [https://CRAN.R-project.org/package=lavaan](https://CRAN.R-project.org/package=lavaan).

Ryan, R. M., & Deci, E. L. (2000). 'Self-determination theory and the facilitation of intrinsic motivation, social development, and well-being', *American Psychologist*, 55(1), 68-78. [https://doi.org/10.1037/0003-066X.55.1.68](https://doi.org/10.1037/0003-066X.55.1.68).

Ryan, R. M., & Deci, E. L. (2017). *Self-determination theory: Basic psychological needs in motivation, development, and wellness*. The Guilford Press. [https://doi.org/10.1521/978.14625/28806](https://doi.org/10.1521/978.14625/28806).

Salvy, S. J., Roemmich, J. N., Bowker, J. C., Romero, N. D., Stadler, P. J., & Epstein, L. H. (2009). 'Effect of peers and friends on youth physical activity and motivation to be physically active', *Journal of Pediatric Psychology*, 34(2), 217-225. [https://doi.org/10.1093/jpepsy/jsn071](https://doi.org/10.1093/jpepsy/jsn071).

Sahin, E., Topkaya, N., & Kurkcu, R. (2016). 'Sex and age differences in achievement goal orientations in Turkish adolescents', *Journal of Education and Practice*, 27(7), 149-156. [https://www.researchgate.net/publication/317721737_Sex_and_Age_Differences_in_Achievement_Goal_Orientations_in_Turkish_Adolescents](https://www.researchgate.net/publication/317721737_Sex_and_Age_Differences_in_Achievement_Goal_Orientations_in_Turkish_Adolescents). Accessed: 2 October 2025.

Shao, T., & Zhou, X. (2023). 'Correlates of physical activity habits in adolescents: A systematic review', *Frontiers in Physiology*, 14, 1131195. [https://doi.org/10.3389/fphys.2023.1131195](https://doi.org/10.3389/fphys.2023.1131195).

Sebire, S. J., Jago, R., Fox, K. R., Edwards, M. J., Thompson, J. L. (2013). 'Testing a self-determination theory model of children's physical activity motivation: a cross-sectional study', *International Journal of Behavioral Nutrition and Physical Activity*, 10, 111. [https://doi.org/10.1186/1479-5868-10-111](https://doi.org/10.1186/1479-5868-10-111).

Sport England. (2024). *Active Lives Children and Young People Survey, 2022-2023* [data collection]. UK Data Service. SN: 9286, DOI: [http://doi.org/10.5255/UKDA-SN-9286-1](http://doi.org/10.5255/UKDA-SN-9286-1). Accessed: 19 August.

Sport England, Ipsos. (2025). *Active Lives Adults Survey, 2022-2023* [data collection]. 2nd Edition. UK Data Service. SN: 9288, DOI: [http://doi.org/10.5255/UKDA-SN-9288-2](http://doi.org/10.5255/UKDA-SN-9288-2). Accessed: 19 August.

Standage, M., Curran, T., & Rouse, P. C. (2019). 'Self-determination-based theories of sport, exercise, and physical activity motivation', in T. S. Horn & A. L. Smith (eds.), *Advances in Sport and Exercise Psychology* (4th Edition), pp. 289-311. Human Kinetics.

Strain, T., Flaxman, S., Guthold, R., Semenova, E., Cowan, M., Riley, L. M., Bull, F. C., Stevens, G. A., & Country Data Author Group (2024). 'National, regional, and global trends in insufficient physical activity among adults from 2000 to 2022: a pooled analysis of 507 population-based surveys with 5·7 million participants', *The Lancet Global Health*, 12(8), e1232-e1243. [https://doi.org/10.1016/S2214-109X(24)00150-5](https://doi.org/10.1016/S2214-109X%2824%2900150-5).

Tapia-Serrano, M. Á., López-Gajardo, M. Á., Sánchez-Miguel, P. A., Llanos-Muñoz, R., & Burgueño, R. (2024). 'Analysis of motivational profiles of physical activity behavior in primary school students: A self-determination theory-based perspective', *Personality and Individual Differences*, 231, 112837. [https://doi.org/10.1016/j.paid.2024.112837](https://doi.org/10.1016/j.paid.2024.112837).

Teixeira, P. J., Carraça, E. V., Markland, D., Silva, M. N., & Ryan, R. M. (2012). 'Exercise, physical activity, and self-determination theory: a systematic review', *International Journal of Behavioral Nutrition and Physical Activity*, 9, 78. [https://doi.org/10.1186/1479-5868-9-78](https://doi.org/10.1186/1479-5868-9-78).

Teuber, M., Leyhr, D., & Sudeck, G. (2024). 'Physical activity improves stress load, recovery, and academic performance-related parameters among university students: a longitudinal study on daily level', *BMC Public Health*, 24, 598. [https://doi.org/10.1186/s12889-024-10598-3](https://doi.org/10.1186/s12889-024-10598-3).

Trigueros, R., Aguilar-Parra, J. M., López-Liria, R., & Rocamora, P. (2019). 'The dark side of the self-determination theory and its influence on the emotional and cognitive processes of students in physical education', *International Journal of Environmental Research and Public Health*, 16(22), 4444. [https://doi.org/10.3390/ijerph16224444](https://doi.org/10.3390/ijerph16224444).

Tsai, T.-H., Wong, A. M., Lee, H.-F., & Tseng, K. C. (2022). 'A study on the motivation of older adults to participate in exercise or physical fitness activities', *Sustainability*, 14(10), 6355. [https://doi.org/10.3390/su14106355](https://doi.org/10.3390/su14106355).

Varma, V. R., Dey, D., Leroux, A., Di, J., Urbanek, J., Xiao, L., & Zipunnikov, V. (2017). 'Re-evaluating the effect of age on physical activity over the lifespan', *Preventive Medicine*, 101, 102-108. [https://doi.org/10.1016/j.ypmed.2017.05.030](https://doi.org/10.1016/j.ypmed.2017.05.030).

Visek, A. J., Achrati, S. M., Mannix, H., McDonnell, K., Harris, B. S., & DiPietro, L. (2015). 'The fun integration theory: toward sustaining children and adolescents sport participation', *Journal of Physical Activity & Health*, 12(3), 424-433. [https://doi.org/10.1123/jpah.2013-0180](https://doi.org/10.1123/jpah.2013-0180).

Vuckovic, V., & Duric, S. (2024). 'Motivational variations in fitness: a population study of exercise modalities, gender and relationship status', *Frontiers in Psychology*, 15, 1377947. [https://doi.org/10.3389/fpsyg.2024.1377947](https://doi.org/10.3389/fpsyg.2024.1377947).

Wang, J., & Syed Ali, S. K. B. (2025). A systematic review of the relationships between attitude, subjective norms, perceived behavioural control, and exercise intention. *The Open Public Health Journal*, *18*(1). [https://doi.org/10.2174/0118749445394456250605043155](https://doi.org/10.2174/0118749445394456250605043155)

Wang, L., & Wang, L. (2015). 'Using theory of planned behavior to predict the physical activity of children: Probing gender differences', *BioMed Research International*, 2015, 536904. [https://doi.org/10.1155/2015/536904](https://doi.org/10.1155/2015/536904).

White, R. L., Vella, S., Biddle, S., White, R. L., Vella, S., Biddle, S., Sutcliffe, J., Guagliano, J. M., Uddin, R., Burgin, A., Apostolopoulos, M., Nguyen, T., Young, C., Taylor, N., Lilley, S., & Teychenne, M. (2024). 'Physical activity and mental health: a systematic review and best-evidence synthesis of mediation and moderation studies', *International Journal of Behavioral Nutrition and Physical Activity*, 21, 134. [https://doi.org/10.1186/s12966-024-01676-6](https://doi.org/10.1186/s12966-024-01676-6).

World Health Organization. (2024). *Nearly 1.8 billion adults at risk of disease from not doing enough physical activity*. [https://www.who.int/news/item/26-06-2024-nearly-1.8-billion-adults-at-risk-of-disease-from-not-doing-enough-physical-activity](https://www.who.int/news/item/26-06-2024-nearly-1.8-billion-adults-at-risk-of-disease-from-not-doing-enough-physical-activity). Accessed: 27 October 2025.

\newpage

# Appendix A - Survey Questions

## Adults

```{r va, table.align='left',tab.cap="Adult survey questions."}
opts <- options(knitr.kable.NA = '')
kable(va,align='l',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```


## Youths

```{r vc, fig.align='center',tab.cap="Youth survey questions."}
opts <- options(knitr.kable.NA = '')
kable(vc,align='l',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

\newpage

# Appendix B - Exercise Types

## Adults

The total minutes of exercise in this dissertation are calculated as the sum of vigorous (original variable: DUR_HVY_CAPPED_SPORTCOUNT_A01) and moderate plus (original variable: DUR_MOD_CAPPED_SPORTCOUNT_A01) activities. Both are capped at 1680 minutes per week.

### Vigorous Exercises

* 11 a-side football
* 13 a-side rugby league
* 15 a-side rugby union
* Bootcamp (e.g. drill sergeant military fitness)
* Boxing
* Boxing class (e.g. Boxercise body combat)
* Cardio class (e.g. aerobics step aerobics body attack)
* Circuit or cross training, cross fit, HIT or boot camp
* Circuit training
* Cross fit
* Cross training
* Cycle class (e.g. spinning RPM)
* Cyclo-cross
* Fell or trail running
* Field hockey
* Futsal
* High intensity (e.g.  HIT insanity)
* Hockey
* Indoor cycling - in a class
* Ju-Jitsu
* Karate
* Mountain biking
* Obstacle course (e.g. Tough Mudder Spartan Rat Race)
* Parkour or free running
* Road cycling or racing
* Rugby union
* Running or jogging
* Small sided football
* Squash or racketball
* Taekwondo
* Track and field athletics
* Triathlon (includes aquathlon and duathlon)
* Weightlifting or powerlifting (using a barbell)


### Moderate Exercises

* A session combining several gym or fitness machines or activities
* Aikido
* Badminton
* Baseball or softball
* Basketball
* BMX
* Body weight exercises (e.g. pull ups press ups sit ups)
* Cheerleading
* Chinese martial arts
* Climbing or bouldering wall
* Climbing or mountaineering
* Cross training machine (e.g. Cross trainer SkiErg)
* Cycling for leisure
* Cycling for leisure
* Cycling for Leisure and all other cycling
* Cycling for travel (including commuting)
* Cycling for travel incl commuting
* Dance-based class (e.g. Zumba fitsteps ravercise or body jam)
* Dance-based class (e.g. Zumba, fitsteps, ravercise or body jam) 
* Dressage
* Eventing
* Exercise bike
* Exercise machine
* Football
* Free weights (includes kettlebells and dumb-bells)
* Gymnastics
* Gymnastics or trampolining
* Handball
* Hill and mountain walking, hiking, mountaineering
* Hill or mountain walking or hiking
* Hill or mountain walking or hiking
* Indoor cycling - not in a class
* Judo
* Lacrosse
* Martial arts
* Mountaineering and scrambling
* Netball
* Other exercise machine
* Other football
* Other horse riding
* Resistance weights machines
* Rock climbing or bouldering
* Rounders
* Rounders
* Rowing
* Rowing machine
* Rowing machine
* Rugby league
* Rugby sevens
* Running machine or treadmill
* Schooling
* Show jumping
* Skiing
* Skiing or snowboarding
* Skipping
* Snowboarding
* Step machine
* Surfing, body surfing or body boarding
* Tag or other rugby league
* Tag or other rugby union
* Tennis
* Touch rugby
* Touch rugby league
* Touch rugby league
* Touch rugby union
* Touch rugby union
* Track cycling
* Trampolining
* Treadmill
* Volleyball
* Walking football
* Water based rowing
* Water polo
* Water polo
* Water-based class (e.g. aquaerobics aquafit)
* Weights (did not specify whether free weights or resistance weights)
* Weights-based class (e.g. body pump kettlebell)
* Wrestling

## Youths

Moderate and vigorous exercises were combined into a single variable (original variable name: mins_modplus_outschool_Week_ALL). According to Sport England (2024), moderate activity involves raising one's heart rate and feeling slightly out of breath, whereas vigorous activity refers to exertion that causes breathlessness or sweating, such that only a few words can be spoken without pausing for breath.

### Moderate and Vigorous Exercises:

* Cycling for fun
* Dancing (include online or TV led e.g. TikTok dances)
* Trampolining (including in a garden, at a trampoline centre, or as part of a club)
* Playing it, tag, chase, sardines or other running game
* Football
* Netball
* Hockey
* Rugby (including tag rugby)
* Touch or tag Rugby
* Contact rugby (rugby union)
* Rugby league (contact)
* Basketball
* Cheerleading
* Running, jogging, cross-country
* Field athletics
* Gym or fitness (fitness/online class e.g., push-ups, sit-ups, yoga, etc or using exercise machines e.g. rowing machine, exercise bike, running machine)
* Judo, karate, taekwondo and other martial arts
* Sports day events
* Boxing


\newpage

# Appendix C - Likert-Scale Response Summary Statistics

```{r adult.summary, fig.align='center',tab.cap="Statistics on Likert-scale responses in adults."}
opts <- options(knitr.kable.NA = '')
kable(list(adult.summary),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

```{r child.summary, fig.align='center',tab.cap="Statistics on Likert-scale responses in youths."}
opts <- options(knitr.kable.NA = '')
kable(list(child.summary),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```


\newpage

# Appendix D - LCA Models and Fit Indices

## Adults

```{r lca.indices.ad1, fig.align='center',tab.cap="LCA fit indices, adults."}
opts <- options(knitr.kable.NA = '')
kable(list(ad.lca.output),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

```{r lca.indices.ad2, fig.align='center',tab.cap="BLRT results, adults."}
opts <- options(knitr.kable.NA = '')
kable(list(blrt.ad$output),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

## Youths

```{r lca.indices.ch1, fig.align='center',tab.cap="LCA fit indices, youths."}
opts <- options(knitr.kable.NA = '')
kable(list(ch.lca.output),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

```{r lca.indices.ch2, fig.align='center',tab.cap="BLRT results, youths."}
opts <- options(knitr.kable.NA = '')
kable(list(blrt.ch$output),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```


\newpage

# Appendix E - R Code (Data Cleaning)

```{r data.R, comment = ""}
source("data1.R", echo = T, print.eval = F,
       max.deparse.length=Inf, keep.source=T)
```

\newpage

# Appendix F - R Code (SEM)
```{r SEM.R, comment = ""}
source("SEM.R", echo = T, print.eval = F,
       max.deparse.length=Inf, keep.source=T)
```

\newpage

# Appendix G - R Code (LCA)
```{r LCA.R, comment = ""}
source("LCA1.R", echo = T, print.eval = F,
       max.deparse.length=Inf, keep.source=T)
```

\newpage

# Appendix H - R Code (Visualization)
```{r Visualization.R, comment = ""}
source("Visualization.R", echo = T, print.eval = F,
       max.deparse.length=Inf, keep.source=T)
```

\newpage

# Appendix I - R code (R Markdown)

```{r get-labels, echo=FALSE, eval=TRUE}
labs = knitr::all_labels()
labs = setdiff(labs, c("setup", "get-labels"))
```

```{r all-code, ref.label=labs, echo=TRUE, eval=FALSE}
```