---
title: "Dissertation"
date: "2025-10-30"
author: "Student ID: 11484265"
output:
  pdf_document: default
  html_document: default
  output: pdf_document
  fontsize: 12pt
  urlcolor: "blue"
  geometry: margin=2in
---

rmdwc::rmdcount('Paper.rmd')

do results and conclusion 
add that social cognitive theory is outdated (if time)
tidy up code

\pagenumbering{arabic}

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo=FALSE, 
                      eval=TRUE, message=FALSE, 
                      warning=FALSE, fig.align='center',
                      fig.dim=c(5,3))
                      
library(rmarkdown)
library(tidyverse)
library(tinytex)
library(checkpoint)
library(knitr)
library(kableExtra)
library(manipulateWidget)

library(scales)
library(lavaan)
library(car)
library(ggplot2)
library(nnet)
library(tidyLPA)
library(poLCA)
library(poLCAExtra)

```


\newpage 

```{r data, include=FALSE, cache=F}
source(file="data.R")
```

```{r SEM, include=FALSE, cache=F}
source(file="SEM.R")
```

```{r LCA, include=FALSE, cache=F}
source(file="LCA.R")
```

```{r Visualization, include=FALSE, cache=F}
source(file="Visualization.R")
```

```{=latex}
\setcounter{tocdepth}{3}
\tableofcontents
```

\newpage 

# Introduction

Physical inactivity remains widespread across different age groups and represents a major public health issue (Mitchell, 2019). Nearly one third of the global population suffers from insufficient physical activity, a 9% increase over the past two decades (Strain et al., 2024). Beyond its physiological consequences, participation in physical activity (PA) is shaped by social, developmental, and motivational factors. Motivation has consistently been identified as a central determinant of behaviour, including PA (Daley & Duda, 2006; Deci & Ryan, 2008; Duncan et al., 2010).

However, most existing studies examine age-related differences within either youths or adults, leaving it unclear whether the same motivational mechanisms operate across life stages. This gap limits the generalisability of findings and the design of effective interventions. To address this, the present study directly compares motivational patterns and their behavioural implications across youths and adults. Latent class analysis is used to identify distinct motivational profiles within each group, while multigroup structural equation modelling tests whether the relationships between motives and physical activity differ by age. On this basis, the study addresses the following research questions:

1. Do perceived exercise motives influence physical activity differently in youths and adults?

2. How do age differences shape dominant exercise motives within youth and adult groups?

# Literature Review

Motivation in physical activity research has been interpreted through several theoretical lenses. Among these, Self-Determination Theory (SDT) and the Theory of Planned Behaviour (TPB) provide complementary perspectives: SDT explains the quality and internalisation of motives, whereas TPB models the intentional processes linking attitudes and perceived control to behaviour.

## Self-Determination Theory (SDT)

Self-Determination Theory (SDT) is a popular framework that has gained popularity in the past two decades. Developed by Deci & Ryan, it has shed light on the understanding of human motivation, emphasizing the degree to which behavior is self-determined versus controlled. Motivation is distinguished in terms of autonomy, with autonomous forms associated with volition and self-endorsement of behavior, and controlled forms reflecting pressure from external contingencies or internalized demands such as guilt or ego involvement. Within SDT, extrinsic motivation is subdivided into external regulation, introjected regulation, identified regulation, and integrated regulation, representing increasing internalization of instrumental behaviors. Intrinsic motivation, by contrast, arises from interest, enjoyment, or inherent satisfaction in the activity itself. These motivational types have implications for physical activity, as more autonomous motives are linked to sustained engagement, positive psychological outcomes, and long-term health (CITE).

In physical activity and health behaviour research, this framework is especially influential because it links the quality of motivation, rather than its quantity, to long-term adherence and wellbeing. Empirical studies confirm that need-supportive interventions can improve exercise uptake, reduce dropout, and enhance both physical and psychological wellbeing (Teixeira et al., 2012; Ng et al., 2012). Consequently, SDT provides not just a descriptive model of motivation but also a prescriptive guide for designing effective physical activity and health promotion programmes.

However, SDT research on physical activity often relies on self-reported motivation and activity measures, which may inflate associations due to shared method variance. Moreover, longitudinal tests of SDT in adulthood remain limited, making developmental generalisations tentative.


## Theory of Planned Behaviour (TPB)

The Theory of Planned Behavior (TPB), proposed by Ajzen in 1991, is a widely used framework for predicting and understanding human behavior, emphasizing the role of intention as the proximal determinant of action. According to TPB, behavioral intentions are influenced by three key factors: attitudes toward the behavior, subjective norms, and perceived behavioral control. Attitudes reflect an individual’s positive or negative evaluation of performing the behavior, subjective norms capture perceived social pressure from significant others to engage or not engage, and perceived behavioral control represents the perceived ease or difficulty of performing the behavior, akin to self-efficacy. These factors interact to shape intention, which in turn predicts behavior, although perceived behavioral control can also have a direct effect on behavior. In the context of physical activity, TPB has been used to explain variations in exercise participation across age groups, demonstrating that stronger intentions—supported by favorable attitudes, positive social norms, and higher perceived control—are associated with higher levels of activity. However, the framework also recognizes that intentions do not always translate into behavior, highlighting the importance of situational constraints and individual capabilities.

TPB has been criticised for its limited explanatory power beyond intention formation, neglecting affective and identity-related processes emphasised in SDT. Meta-analytic evidence shows that TPB variables explain intention well but only moderate proportions of actual behaviour (Rhodes & de Bruijn, 2013).

Together, these frameworks imply that motivation for physical activity operates through both qualitative (autonomous vs. controlled) and structural (attitude, norm, control) mechanisms. From a statistical perspective, this supports using latent class analysis to identify subgroups defined by motivational quality, and structural equation modelling to examine predictive pathways between motives and behaviour.

\newpage 

## Age Differences

Across the lifespan, intrinsic and personally meaningful motivations are consistently associated with higher levels of physical activity. However, intrinsic motivation tends to decline with age (Brunet & Sabiston, 2011; Dishman et al., 2018; de Maio Nascimento, 2023), suggesting that the factors driving engagement shift across developmental stages. Environmental opportunities, such as access to facilities or structured programs, appear to facilitate activity regardless of age, providing a common context that supports engagement across populations.

In youths, intrinsic motivation, enjoyment, and social support are strong predictors of habitual physical activity. Longitudinal evidence indicates that adolescents who maintain higher intrinsic motivation and personally meaningful goals—particularly those emphasizing enjoyment—remain more active over time, whereas declines in social, competence, or appearance goals can weaken engagement (Dishman et al., 2018). Cross-sectional studies further show that higher motivation profiles in adolescents are associated with adaptive outcomes such as responsibility, resilience, and social support (Manzano-Sánchez et al., 2019; Heredia-León et al., 2021). Collectively, these findings suggest that enjoyment- and social-oriented motives play a particularly salient role in sustaining physical activity in younger populations.

Among adults, motivation tends to shift toward health- and fitness-related goals. Adults frequently report exercising for reasons aligned with personal health, fitness, and psychological well-being rather than external rewards (Brunet & Sabiston, 2011; Nascimento et al., 2023). Relaxation and stress relief have also been highlighted as important motivating factors, contributing to voluntary adherence and psychological benefits (de Maio Nascimento, 2021; Vuckovic, 2015; Kilgour, 2005). Evidence suggests that these motives remain relatively stable across younger and older adults, with comparable importance reported by women over 50 (Kilgour, 2005; Vuckovic, 2015). Overall, fitness- and relaxation-related motives appear more prominent in adults than in youths.

While specific motives vary by age, some contextual factors influence physical activity similarly across groups. Access to facilities, opportunities to participate, and supportive environments facilitate engagement across the lifespan, providing a baseline influence independent of age-specific motivational differences (Brunet & Sabiston, 2011; Dishman et al., 2018).

These patterns suggest clear age-related differences in the influence of specific motives. Enjoyment, social, and guilt motives are especially relevant for youths, whereas fitness- and relaxation-related motives are more salient in adults. Environmental opportunities appear to exert a similar influence across age groups. However, much of this evidence is cross-sectional, making it difficult to determine whether observed differences reflect age effects or cohort effects. Moreover, the relative salience of enjoyment versus health motives may depend on contextual factors such as school environments or occupational demands, which are rarely controlled for.

Hypothesis 1: The influence of exercise motives on physical activity differs between youths and adults.

* H1(a). Enjoyment, social, and guilt motives are more influential in youths than in adults.

* H1(b). Fitness and relaxation motives are more influential motives in adults than in youths.

* H1(c). Environmental opportunities influence physical activity similarly across age groups.

## Motivational Profiles


Hypothesis 2: Distinct motivational profiles exist within youths and within adults.

In adults, Ostendorf et al. (2021) identified three primary motivational profiles among individuals with overweight or obesity: high autonomous, high combined, and moderate combined. The high autonomous profile was characterized by strong intrinsic and identified motivations, with minimal influence of external or introjected regulation. The high combined profile reflected elevated levels across all regulatory types, while the moderate combined profile exhibited intermediate levels on all regulations. Longitudinally, the high autonomous profile demonstrated the least decline in moderate-to-vigorous physical activity during transitions from supervised to unsupervised exercise, suggesting that intrinsic and identified motivations support sustained behavior, whereas moderate-to-high external regulation may require additional support for continued engagement. Nuss et al. (2023) further identified four motivational profiles in Canadian adults, showing that combinations of controlled and autonomous motivation may synergistically support activity, with low overall motivation corresponding to minimal engagement. Collectively, these findings highlight that adult populations are not homogeneous in their motivational profiles, and that the prevalence of lower-engagement profiles increases with age. This supports the hypothesis that, with increasing age, adults are progressively more likely to belong to classes characterized by lower overall agreement across motivational items, with the oldest adults showing the sharpest decline.

In adolescents, Moreno-Murcia et al. (2011) found two primary motivational profiles in physical education students: a self-determined profile, with high intrinsic and identified motivation, and a non-self-determined profile, with elevated external, introjected, and amotivated scores. The self-determined profile was positively associated with Theory of Planned Behavior constructs such as intention, subjective norm, perceived behavioral control, and attitude, indicating that higher autonomous motivation supports favorable cognitions and participation behaviors. Cross-sectional studies further corroborate these trends. Manzano-Sánchez et al. (2019) identified higher motivation profiles among adolescents aged 12–16, which were associated with adaptive outcomes such as responsibility, resilience, and perceived social support. Heredia-León et al. (2021) reported that students with high-quality and high-quantity motivational profiles demonstrated greater intention to be physically active and enjoyment in PE classes, whereas low-quality profiles corresponded to higher boredom and lower engagement. Tapia-Serrano et al. (2022) identified five profiles in children, ranging from highly amotivated to autonomously motivated, showing that even controlled motivation can promote short-term engagement, though autonomy supports more sustained activity. While these studies collectively demonstrate heterogeneity in motivational orientation, they differ in sample characteristics, activity settings, and profile extraction methods, complicating direct comparison across age groups. Few have tested whether the same latent structures hold equivalently in youths and adults, underscoring the need for multigroup analysis. Nevertheless, younger populations exhibit a range of motivational profiles, and the likelihood of belonging to profiles with lower motivation increases with age. This aligns with the hypothesis that, as age increases, youths are progressively more likely to belong to classes characterized by lower overall agreement across motivational items.

* H2(a). With increasing age, the likelihood of adults belonging to classes with lower overall agreement across motivational items increases, reflecting a gradual decline in motivation, with the oldest adults showing the steepest decrease.

* H2(b). With increasing age, the likelihood of youths belonging to classes with lower overall agreement across motivational items increases, reflecting a gradual decline in motivation.

Existing research indicates that physical activity motivation varies systematically across the lifespan, yet most studies examine either youths or adults in isolation, use cross-sectional designs, and apply different measurement structures. The current study extends prior work by jointly modelling motivational constructs and their behavioural implications across age groups, employing latent class analysis to identify subgroups of motivation and multigroup SEM to test predictive equivalence.

\newpage 

# Data and Methods

## Data

The analyses draw on large-scale survey data collected by Ipsos on behalf of Sport England (2024, 2025). These datasets were chosen because the youth and adult surveys share a parallel structure, with several items identically worded across both instruments. This allows direct comparison of motivational mechanisms across life stages. All motive items are self-reported, single-item measures. Appendix A lists the corresponding survey questions.

Descriptive statistics, bivariate correlations, and variance inflation factors (VIF) were computed for all motive variables to assess distributional properties and multicollinearity. Challenge and relaxation were excluded from the structural equation models (SEM) due to non-equivalent wording between the youth and adult questionnaires but retained for latent class analysis (LCA).

The SEM analyses included `r prettyNum(nrow(adult.bi), big.mark=",")` adult and `r prettyNum(nrow(child.bi), big.mark=",")` youth respondents; the LCA included `r prettyNum(nrow(adult.lik), big.mark=",")` and `r prettyNum(nrow(child.lik), big.mark=",")`, respectively . To maintain comparability and avoid skew from small subgroups, only cisgender adults without disabilities were retained. Ethnicity and gender were included as control variables. Observations with missing data on relevant items were likewise deleted.

Motivational and behavioural variables:

* Enjoyment – finding exercise satisfying or pleasurable.

* Social – exercising for fun or connection with friends.

* Fitness – exercising to maintain or improve physical health.

* Guilt – exercising out of obligation or self-pressure.

* Opportunity – having access, time, or suitable conditions to exercise.

* Relaxation – exercising to relieve stress or worry.

* Ability – self-perceived competence and confidence in physical activity.

* Importance – recognising the personal value or significance of physical activity.

* Challenge – persistence and enjoyment in pursuing difficult tasks.

* Minutes – total weekly minutes of moderate-to-vigorous physical activity (see Appendix B for included activities).


\newpage 

### Adult Dataset

The adult data derive from the Active Lives Adult Survey (Year 8: 2022–2023), conducted by Ipsos for Sport England with additional support from the Office for Health Improvement and Disparities (OHID). The survey is designed to provide a nationally representative picture of adult participation in sport and physical activity in England.

Sampling was based on the Postcode Address File (PAF), using a random probability design. Up to two adults (aged 16+) per selected household were invited to participate. A push-to-web approach was employed: two initial mailings invited online completion, followed by a third mailing with a paper questionnaire to increase response rates, and a final reminder letter. Data collection occurred from November 2022 to November 2023, with approximately 500 responses targeted per local authority to minimise seasonal and geographic bias.

A total of 173,950 adults completed the survey. Weighting adjustments were applied to correct for differential response probabilities. The questionnaire covered a broad range of topics, including physical activity frequency and duration, sports participation, volunteering, club membership, and motivational factors, along with demographic variables such as age, gender, education, and socio-economic status.


```{r gg.ad.demo1, out.width='32%',fig.show='hold',fig.align='center', fig.cap='Distribution of age, gender, and ethnicity in adults.'}
gg.ad.age
gg.ad.gend
gg.ad.eth
```

Gender and ethnicity distributions were highly similar across the subsamples used for the SEM and LCA analyses. Females comprised `r prettyNum(prop.table(table(adult.bi$gender))[[2]]*100, big.mark=",", digits = 2)`% of the sample, and White British participants accounted for `r prettyNum(prop.table(table(adult.lik$eth))[[1]]*100, big.mark=",", digits = 2)`%. The youngest and oldest age groups were collapsed due to small cell sizes at the extremes. Specifically, adults aged 64–75 and 76+ were combined, as preliminary analyses indicated no meaningful differences in their motivational profiles. Collapsing these categories improved model stability without obscuring theoretically relevant distinctions, which are less pronounced in later adulthood. After adjustment, the age distribution remained moderately skewed but acceptable for analysis. Cases with missing values on any motivational variables were excluded from all models.

```{r adult.summary.bi, fig.align='center',tab.cap="Proportion of positive responses in adults."}
opts <- options(knitr.kable.NA = '')
kable(list(adult.summary.bi),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

The distributions of positive responses exhibit a left skew, suggesting that participants rarely chose the highest agreement category for most items.


```{r vars.ad,fig.height= '150%',fig.show='hold',fig.align='center', fig.cap='Distribution of adult responses for each motive on the Likert scale.'}

gg.vars.ad

```

Each item was rated on a five-point Likert scale ranging from 1 (strongly agree) to 5 (strongly disagree). Due to low frequencies in the higher disagreement categories, responses of 4 (disagree) and 5 (strongly disagree) were combined to reduce sparsity and minimize distortion.

For most motive items, the proportion of adults expressing strong positive agreement declines with age. Older adults are less likely than younger adults to strongly endorse beliefs such as feeling capable of engaging in physical activity, valuing it for the challenge, or pursuing it for enjoyment and fitness. They are also less likely to report strong guilt for not exercising, to perceive sufficient opportunities for activity, or to consider physical activity an important means of relaxation. Additional descriptive statistics, including mean, median, standard deviation, and the percentage of missing values (excluded from the final dataset), are provided in Appendix C.

\newpage 

### Youth Dataset

The youth dataset used in this dissertation is drawn from the Active Lives Children and Young People Survey (Year 6: 2022–2023), conducted by Ipsos on behalf of Sport England. It is a large-scale, school-based online survey administered to pupils in Years 1–11, their parents (for Years 1–2 pupils), and teachers. Only pupil responses from year 6-11 (age 11-16) were used in this dissertation. The pupil questionnaire focused on participation in sport and physical activity over the previous week, alongside items on swimming, cycling, volunteering, wellbeing, and attitudes towards physical activity. It also included classification questions such as gender, disability, and long-term health conditions. 

The sampling strategy was designed to permit analysis at both national and local authority levels. A stratified three-stage sampling process was used: schools were first sampled from the January 2021 school census, then three year groups were randomly selected within each participating school, and finally, one mixed-ability class was chosen per selected year group. Fieldwork was carried out in three phases aligned with the academic terms (September 2022–July 2023). Pupils typically completed the survey at school under teacher supervision, although in some secondary schools it could also be set as homework. To encourage participation, schools received credits for sports equipment and, if response thresholds were met, school-level feedback reports. 

```{r gg.ch.demo, out.width='32%',fig.show='hold',fig.align='center', fig.cap='Distribution of age, gender, and ethnicity in youths.'}

gg.ch.age
gg.ch.gend
gg.ch.eth

```

`r prettyNum(prop.table(table(child.lik$gender))[[2]]*100, big.mark=",", digits = 2)`% of the youth sample are female, and `r prettyNum(prop.table(table(child.lik$eth))[[1]]*100, big.mark=",", digits = 2)`% are White British. Most participants are aged 12–15, with only 4.63% aged 16. Although 16-year-olds represent a small proportion of the sample, they were not collapsed with the 15-year-olds. This decision was guided by substantive, rather than purely numerical, considerations: the transition from 15 to 16 marks a distinct developmental and social stage—such as reaching legal thresholds, completing compulsory schooling, and gaining increased autonomy—which may influence physical activity motives. Retaining this distinction allows examination of whether these developmental turning points are reflected in motivational patterns, even within a relatively small subgroup.

```{r child.summary.bi, fig.align='center',tab.cap="Proportion of positive responses in youths."}
opts <- options(knitr.kable.NA = '')
kable(list(child.summary.bi),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

The distributions of positive responses exhibit a left skew, similar to the adults, suggesting that participants rarely chose the highest agreement category across items.

```{r vars.ch,,fig.height= '130%', fig.show='hold',fig.align='center', fig.cap='Distribution of youth responses for each motive on the Likert scale.'}

gg.vars.ch

```


Each item was rated on a four-point Likert scale ranging from 1 (strongly agree) to 4 (strongly disagree). As with the adult dataset, responses of 3 (disagree) and 4 (strongly disagree) were combined due to low frequencies. Missing responses and “I don’t know” answers were excluded from the dataset.

After collapsing, the items are generally balanced, except for social, guilt, and relaxation. Response patterns among youths are less consistent than those observed in adults. With increasing age, youths are less likely to report exercising for challenge or social reasons, while a larger proportion indicate exercising due to guilt or for relaxation. Among the oldest youth group, there is a slight but noticeable increase in the proportion who “strongly agree” that ability and fitness are motivating factors. Across all ages, only a small minority report lacking understanding of the importance of physical activity. Additional descriptive statistics, including means, medians, standard deviations, and percentages of missing values (excluded from analysis), are provided in Appendix C.

\newpage 

## Multigroup Structural Equation Modeling (SEM)

Differences in the relationships between self-reported motives and PA levels across youths and adults were examined while controlling for demographic factors (see Appendix E for code and model specifications).

Motivation variables included enjoyment, social, fitness, guilt, and opportunity. To account for differences in Likert scales between adults and youths, all motivation variables were dichotomized into “strongly agree” versus “not strongly agree.” Demographic covariates included gender, age, and ethnicity. Gender was limited to female and male due to small sample sizes in other categories, and ethnicity was collapsed into White British and Non-White British. Youth participants included only those aged 11 and older who responded to the relevant items. Adults were grouped into age ranges (16–34, 35–44, 45–54, 55–64, 65–74, 75+) because exact ages were unavailable. The youngest and oldest two adult groups were further collapsed to reduce skew and ensure more balanced distributions. A cap of 1,680 minutes per week was applied to reported PA to minimize the influence of extreme values or potential data entry errors.

Multigroup SEM was used to assess how each motive predicts weekly minutes of PA, allowing direct comparison of pathway strengths between youths and adults. A freely estimated model was first fitted and then compared to constrained models, in which individual or all motive pathways were fixed to equality across groups. This approach enabled evaluation of whether the effects of specific motives differed by age group. Differences in the predictive strength of each motive on PA minutes were also calculated. 

Prior to fitting SEM models, assumptions were assessed. Moderate correlations were observed among the motivational predictors in both youth and adult samples, with most values ranging between 0.2 and 0.5. The highest correlations were between enjoyment and ability in youths (r = 0.656) and between enjoyment and perceived importance in adults (r = 0.674), slightly exceeding the conventional threshold of r = 0.6 for concern about multicollinearity. Standardized residual correlations were also examined to evaluate local fit for both youth and adult SEMs. Most residuals were below 0.35, indicating generally acceptable correspondence between observed and model-implied correlations. The largest residuals occurred among closely related motivational items (enjoyment, fitness, opportunity, and relaxation), with several exceeding 0.4, reflecting misfit due to conceptual overlap and the use of dichotomized indicators. Residuals involving demographics and physical activity minutes were consistently low. These localized discrepancies do not affect the primary purpose of testing whether the strengths of associations between motives and physical activity differ across groups.

\newpage 

## Latent Profile Analysis (LCA)

Latent class analysis (LCA) was conducted separately for youths and adults to explore age-related differences in motivational profiles, using the same set of motivation variables included in the SEM, with the addition of two items with slightly different wording.

The original Likert-scale responses were retained. Additional items capturing similar motivational constructs but worded differently were included (see Appendix B). Motivation variables served as indicators, while ethnicity, gender, age, and education (for adults only) were included as covariates. Ten random starts were used per class model to ensure solution stability.

The optimal number of classes was determined based on BIC elbow plots, relative entropy, bootstrap Vuong-Lo-Mendell-Rubin likelihood ratio tests (BLRT), class proportions, and substantive interpretability. Class-specific statistics were calculated, and multinomial logistic regression was used to assess the effect of age on class membership. Odds ratios and 95% confidence intervals were derived by exponentiating the estimated coefficients and their standard errors:

 (\[\text{OR} = \exp(\hat{\beta}), \quad 95\%\, CI = \exp \big( \hat{\beta} \pm 1.96 \times SE \big) \])

Conditional independence among PA-related items was assessed using standardized bivariate chi-square statistics. However, extremely sparse cells (<1% of observations) inflated these values, making them unreliable indicators of local dependence. Additionally, the key motivational items represent conceptually distinct constructs, so minor violations of conditional independence are less concerning than they would be for highly overlapping indicators. Instead, model evaluation relied on conventional LCA fit indices, including BIC and entropy, alongside substantive interpretability. These criteria indicated that the identified profiles were robust and meaningful despite the limitations of the conditional independence tests.


## Software and Packages (CHECK

All analyses were conducted in R (version 4.4.3). Multigroup structural equation modeling was performed using the lavaan package (Rosseel, 2012), which allows for flexible specification of path models, multigroup comparisons, and model fit evaluation. Latent class analysis (LCA) was conducted separately for youths and adults using poLCA (Linzer & Lewis, 2011) and poLCAExtra (Choi, 2023), providing tools for estimation of class probabilities, fit statistics, and post-hoc regression of class membership on covariates.


\newpage 

# Results
  
• Present the results of the analysis.
• Try to focus on how the results answer your research questions and hypotheses.
• Try to focus on substantive interpretation of the results (and not just if something is significant or not). Are the effects large? Are they substantively important?


## SEM

There are minor yet significant differences in the impact of every motive. 

```{r slopes.diff, fig.align='center',tab.cap="Estimated slopes of each motivational factor in the SEM, with differences calculated as youth minus adult values."}
opts <- options(knitr.kable.NA = '')
kable(list(slopes.diff),align='c',booktabs = T) %>% 
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

All motivational items in the model influenced physical activity in the same direction for youths and adults. Since the variables are binary (1 = strongly agree, 0 = not strongly agree), the coefficients represent the estimated difference in minutes of activity between participants who strongly endorse a motive and those who do not. While all differences between age groups were statistically significant, effect sizes were generally modest, except for opportunity, which had a greater impact in adults. Other motives, including enjoyment, guilt, fitness, and social factors, exhibited smaller differences, indicating broadly similar motivational mechanisms across age groups.

Youths were most strongly influenced by enjoyment and guilt, highlighting the role of intrinsic and internalized motivation in shaping adolescent activity. Adults, by contrast, responded more strongly to opportunity, fitness, and social motives, underscoring the importance of external facilitation, health awareness, and social engagement. 

From an SDT perspective, these findings align with the framework’s distinction between intrinsic, internalized, and external regulation. Youths’ strong responses to enjoyment and guilt/obligation reflect intrinsic and internalized motivations, emphasizing autonomy and internalized norms as key drivers. Adults’ stronger responsiveness to opportunity, fitness, and social motives reflects identified and external regulation, where structured opportunities, competence-related goals, and social reinforcement play a larger role. 

Within the TPB framework, behavior is determined by intention, which is influenced by attitudes, subjective norms, and perceived behavioral control (PBC). For youths, full endorsement of enjoyment and guilt/obligation likely reflects favorable attitudes and perceived social norms supporting activity. For adults, stronger effects of opportunity, fitness, and social motives suggest that PBC, shaped by access to resources and social support, is a key driver of activity. 

H1(a) is partially correct. The social aspect is more prominent in adults.
* H1(a). Enjoyment, social, and guilt motives are more influential in youths than in adults.

H1(b) is correct. Will check the second part of hypothesis in lca.
(add in paragraph in intro, Marcelo de Maio Nascimento, Vojko Vuckovic, Alixe H M Kilgour)
* H1(b). Fitness and relaxation are more influential motives in adults than in youths, particularly in middle-aged and older groups.


Actually opportunity affects adults a lot more :O
* H1(c). Environmental opportunities influence physical activity similarly across age groups.




\newpage 

## LCA

Median mins per class, rather than mean due to skew in distribution of minutes exercised (right skew), is calculated. 

### Adults

```{r elbow.llik.ad, out.width='49%',fig.show='hold',fig.align='center', fig.cap='Elbow plot showing AIC and BIC values, and maximum likelihood plot, across different numbers of latent classes in adults.'}

gg.elbow.ad
gg.llik.ad

```

ad.lca.output?


3 to 6 classes all have relative high likelihood and low AIC and BIC, and average posterior entropy per class. From the elbow plot, 4 classes seem to be the number of classes that start displaying diminishing return. However, the 3-class model has significantly higher relative entropy, and the four-class solution produced two classes with highly similar endorsement patterns, effectively splitting the moderate group without adding meaningful differentiation. The three-class model provides a clearer, more parsimonious representation of motivational profiles while still capturing the small but conceptually important subgroup with generally negative attitudes.


```{r distr.ad, fig.show='hold',fig.align='center', fig.cap='Frequency of adult responses for each motive by class membership.'}

plot(LCAE.ad, nclass = 2)

```


#### Class 1

This class shows high probabilities of strongly agreeing with enjoyment, fitness, ability, opportunity, and importance. Members also moderately endorsed social, challenge, and guilt motives, and showed relatively high agreement with relaxation. Overall, this class represents adults who consistently report strong agreement with a broad range of motivations. This class is labelled the High Engagement class.

#### Class 2

This class is characterized by very low probabilities of strongly agreeing with enjoyment and ability, and by a higher prevalence of neutral or negative responses across most items. Members of this class express the least favorable attitudes toward PA. This class is labelled the Low Engagement class. While this class represents a relatively small proportion of the sample (~6%), it aligns with the raw data, in which very few respondents selected negative responses.

#### Class 3

This class has a mixed attitude toward many motives, with moderate positive endorsement toward enjoyment, fitness, opportunity, ability and importance It represents adults with more ambivalent responses. This class is labelled the Mixed Motivation class.


```{r mins.ad, out.width='49%', fig.show='hold',fig.align='center', fig.cap='Weighted median and interquartile range of minutes exercised per week for adults by class.'}

gg.mins.ad

```
 
Analysis of weekly minutes of physical activity across the adult latent classes highlights substantial differences in exercise behavior. The High Engagement class exhibits the highest levels, with a weighted median of 445 minutes per week and a wide interquartile range of 180 to 840 minutes, indicating considerable variability even among this highly active group. The Mixed Motivation class reports intermediate activity, with a median of 182.5 minutes per week and an IQR of 30 to 457.5 minutes, reflecting both moderate overall engagement and substantial individual differences. In contrast, the Low Engagement class shows markedly lower activity, with a median of only 45 minutes per week and an IQR from 0 to 225.25 minutes, highlighting that while most members are minimally active, some still achieve higher levels. Overall, the broad IQRs across all classes indicate that even within latent classes defined by motivation, actual exercise behavior varies considerably.

```{r byage.ad, out.width='49%',fig.show='hold',fig.align='center', fig.cap='Proportion of adults in each class across age groups.'}

gg.byage.ad

```


The distribution of age groups across the three latent classes highlights notable differences in motivational profiles. The High Engagement class is consistently the second most populous across age groups, though its representation declines gradually with increasing age. The Low Engagement class remains the smallest in all age groups, representing a relatively rare subgroup with low endorsement of motives, and its relative size is largely stable across ages. In contrast, the Mixed Motivation class contains the largest number of adults in every age group, with membership increasing steadily as age rises, suggesting that some individuals may shift from high engagement to mixed motivation over time.

```{r or.ad, fig.align='center',tab.cap="Odds ratios for adult class membership across age groups, with the youngest group (16–34 years) as the reference."}
opts <- options(knitr.kable.NA = '')
kable(list(or.ad),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

Multinomial logistic regression examined the association between age and motivation profile, with the High Engagement class as reference. Examination of the odds of latent class membership across adult age groups, using the youngest group (16–34 years) as the reference, reveals clear age-related patterns. Adults in the youngest age group are over 6.6 times more likely to belong to the High Engagement class than to the Low Engagement class. Conversely, a slightly higher proportion of young adults belong to the Mixed Motivation class, which may partly reflect the overall larger size of this class. For the Low Engagement class, older age groups generally show modestly higher odds of membership compared with the reference, though the differences for age groups 35–44 and 55–64 are not statistically significant. The highest odds are observed in the oldest age group (75+), suggesting that adults in this group are 1.5 times more likely to belong to the Low Engagement class than younger adults. In the Mixed Motivation class, the odds of membership increase consistently with age, with adults aged 65–74 being 1.45 times, and those 75+ being 1.93 times more likely than 16–34-year-olds to belong to this class rather than High Engagement. 

These results support H2(a), which posited that the likelihood of adults belonging to classes with lower overall agreement across motivational items increases with age. The observed shift in class membership with age suggests a substantive decline in overall motivation among adults, particularly in the oldest age groups. This pattern implies that interventions targeting physical activity may need to be tailored to address the motivational needs of older adults, who are disproportionately represented in classes with lower agreement across motivational items. The steep decline among the 75+ group indicates that standard approaches effective for younger adults may be insufficient for maintaining engagement in later life. Conceptually, these findings support the idea that motivation is not static across adulthood; rather, it gradually diminishes, with older adults potentially requiring additional support, reinforcement, or environmental facilitation to sustain participation. In practical terms, understanding these shifts can inform program design, policy decisions, and public health strategies aimed at reducing age-related decreases in activity and engagement.



\newpage 

### Youths

BIC plot indicates 3 or four 4 to reduce BIC the most. BLRT preferred 4 classes. Relative entropy values are `r LCAE.ch$output[2,6]` and `r LCAE.ch$output[3,6]` for 3 and 4 classes, respectively, which shows 3 to have slightly better separation between classes.
The likelihood is similar between 3 and 4 classes. Average posterior probabilities are better in the 3-class model, as all classes have a >.80 pp.


```{r elbow.llik.ch, out.width='49%',fig.show='hold',fig.align='center', fig.cap='Elbow plot showing AIC and BIC values, and maximum likelihood plot, across different numbers of latent classes in youths.'}

gg.elbow.ch
gg.llik.ch

```


Both 3 and 4- class models show promise.  BIC plot does not show a typical elbow shape, as adding more classes beyond 4 actually diminished the fit. This is possibly due to the log-likelihood not increasing significantly with more classes, and BIC's penalty for higher complexity outweighs the improvement

However, the 4-class model contains 2 very similar classes. Hence the 3-class model was chosen.

```{r distr.ch, out.width='100%',fig.show='hold',fig.align='center', fig.cap='Frequency of adult responses for each motive by class membership.'}

plot(LCAE.ch, nclass = 2)

```

#### Class 1

This class exhibits very high probabilities of strongly agreeing with enjoyment, fitness, opportunity, importance, and ability. Endorsement of social, challenge, and relaxation motives is somewhat more moderate, though still leaning toward agreement, and guilt shows a more even split across response options. This group is characterized by consistently positive attitudes toward a wide range of motives. This class will be labelled the High Engagement class.

#### Class 2

This class shows very low probabilities of strongly agreeing with all motives, and negative attitudes toward social, guilt, relaxation, and ability motives. This class will be labelled the Low Engagement class.

#### Class 3

This class primarily endorsed “agree” across most motives, with the exception of guilt. Disagreement is relatively uncommon, but strong agreement is also less prevalent than in the High Engagement class. This class falls between the other two, representing generally positive but not strongly emphatic motivation toward PA. This class will be labelled the Moderate Engagement class.

```{r mins.ch, out.width='100%', fig.show='hold',fig.align='center', fig.cap='Weighted median and interquartile range of minutes exercised per week for youths by class.'}

gg.mins.ch

```

Analysis of weekly minutes of physical activity across the youth latent classes reveals meaningful differences in exercise behavior. The High Engagement class reports the highest levels, with a weighted median of 405 minutes per week and an interquartile range of 185 to 745 minutes, indicating substantial variability even among the most motivated youths. The Mixed Motivation class shows intermediate activity, with a median of 235 minutes per week and an IQR from 90 to 455 minutes, reflecting moderate engagement with considerable individual differences. The Low Engagement class exhibits the lowest activity, with a median of 125 minutes per week and an interquartile range of 30 to 300 minutes, suggesting that while most members are less active, some still achieve moderate levels. These patterns imply that, similar to adults, class membership captures meaningful distinctions in motivation that are associated with actual physical activity.

```{r byage.ch, out.width='100%',fig.show='hold',fig.align='center', fig.cap='Proportion of youths in each class by age.'}

gg.byage.ch

```

The distribution of age groups across the three latent classes highlights notable differences in motivational profiles among youths. The High Engagement and Mixed Motivation classes are roughly similar in size, with the High Engagement class generally decreasing across age groups but showing a sudden increase at age 16. The Low Engagement class is smaller than the other two but not as rare as in the adult sample; it increases steadily with age, with a slight dip at age 16. These patterns suggest that while high engagement may decline during adolescence, a substantial portion of youths maintain strong or mixed motivation, and low engagement gradually becomes more common over time. The sudden increase in High Engagement at age 16 indicates that some youths may experience a late boost in motivation during this stage of adolescence.


```{r or.ch, fig.align='center',tab.cap="Odds ratios for youth class membership across ages, with 11-year-olds as the reference."}
opts <- options(knitr.kable.NA = '')
kable(list(or.ch),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

Examination of the odds of latent class membership across youth age groups, using 11-year-olds as the reference, reveals distinct developmental patterns. At age 11, youths are approximately 5 times less likely to belong to the Low Engagement class compared with High Engagement, whereas they are about equal in odds to belong to the Mixed Motivation class. For the Low Engagement class, the odds of membership increase steadily with age, with 12-year-olds about 1.2 times more likely, 13-year-olds 1.6 times more likely, 14-year-olds 1.7 times more likely, and 15-year-olds 2 times more likely than 11-year-olds to belong to this class, before slightly declining at age 16 (1.4 times more likely). In the Mixed Motivation class, the odds of membership increase modestly from ages 12–15, ranging from 1.0 to 1.2 times more likely than 11-year-olds, before declining slightly at age 16 (0.9 times as likely). These patterns indicate that as youths age, they increasingly shift from High Engagement toward Low Engagement or Mixed Motivation, with a subtle rise in motivation at age 16 suggesting that some youths regain or consolidate engagement in physical activity.

The trends indicate a gradual decline in motivation throughout adolescence, with older youths more frequently represented in classes showing moderate or lower endorsement of motivational items, aligning with H2(b). This pattern suggests that interventions aimed at promoting physical activity in younger populations may need to account for declining motivation with age, emphasizing strategies that sustain engagement through the middle to late teen years. The relative stability in the youngest ages followed by a shift around 15–16 years highlights a critical period for reinforcing positive attitudes toward physical activity. Conceptually, these results indicate that motivation in youths is dynamic and susceptible to age-related changes, reinforcing the importance of early and sustained intervention to maintain high levels of agreement across motivational factors.

\newpage 

# Conclusions

• Summarise what you have found. Restate your questions and hypotheses and show
how you answer them.
• Discuss possible limitations and implications they might have for the results.


## Limitations

adult data
Push-to-web design: Reliance on online completion may underrepresent groups with lower digital literacy or internet access (e.g., older adults, lower-income households), even though a paper option was provided at the third mailing. Response bias: Out of over 560,000 households invited, only ~173,000 valid responses were obtained. This implies a relatively low response rate, raising the possibility that respondents are more health-conscious or sport-engaged than non-respondents. Household-level sampling: Allowing up to two respondents per household could introduce intra-household clustering, potentially inflating similarities between responses. Mode effects: Differences between online and paper responses could influence reported activity levels or willingness to disclose sensitive demographic details. Split questionnaire design: Some questions were only asked to subsets of participants, which led to the exclusion of data from participants who did not answer key questions required for this analysis.


youths data
Self-report data: Pupils’ answers, especially younger ones, may be influenced by recall issues, comprehension, or social desirability (e.g., overreporting activity). Mode of administration: Conducted in schools, often under teacher supervision, which may affect how candid pupils are in reporting behaviours or wellbeing. 



No factor analysis
Wording slightly different between datasets for 2 vars 
Correlation moderate in sem
interpreattion may be diff despite same 
skewed data (like most self-reports
did not log transform minutes, rather just took the median, results are extremely similar and does not detract from comparison
removed "does not know" from youths rather than using it as neutral answer
only counted moderate to heavy exercises, but for older adults they may only be able to perform and benefit equally from light exercises

demogrpahics like gender, ethnicity, disability, relationship, not considered

motive items limited to only the ones that are worded the same or extremely similar
data is cross sectinoal

• Discuss implications for theory and/or policy based on what you found.


\newpage 

# References

Sport England. (2024). Active Lives Children and Young People Survey, 2022-2023. [data
collection]. UK Data Service. SN: 9286, DOI: http://doi.org/10.5255/UKDA-SN-9286-1

Sport England, Ipsos. (2025). Active Lives Adults Survey, 2022-2023. [data collection].
2nd Edition. UK Data Service. SN: 9288, DOI: http://doi.org/10.5255/UKDA-SN-9288-2

Vuckovic, V., & Duric, S. (2024). Motivational variations in fitness: a population study of exercise modalities, gender and relationship status. Frontiers in psychology, 15, 1377947. https://doi.org/10.3389/fpsyg.2024.1377947

Ostendorf, D.M., Schmiege, S.J., Conroy, D.E. et al. Motivational profiles and change in physical activity during a weight loss intervention: a secondary data analysis. Int J Behav Nutr Phys Act 18, 158 (2021). https://doi.org/10.1186/s12966-021-01225-5

Moreno-Murcia, J. A., Cervelló Gimeno, E., Hernández, E. H., Pedreño, N. B., & Rodríguez Marín, J. J. (2013). Motivational profiles in physical education and their relation to the theory of planned behavior. Journal of sports science & medicine, 12(3), 551–558. 

Nuss, K., Sui, W., Rhodes, R., & Liu, S. (2023). Motivational Profiles and Associations With Physical Activity Before, During, and After the COVID-19 Pandemic: Retrospective Study. JMIR formative research, 7, e43411. https://doi.org/10.2196/43411

Ajzen, Icek (1991). "The theory of planned behavior". Organizational Behavior and Human Decision Processes. 50 (2): 179–211. doi:10.1016/0749-5978(91)90020-T. hdl:11323/1038. S2CID 260959149.

Dishman, R. K., McIver, K. L., Dowda, M., & Pate, R. R. (2018). Declining Physical Activity and Motivation from Middle School to High School. Medicine and science in sports and exercise, 50(6), 1206–1215. https://doi.org/10.1249/MSS.0000000000001542

Varma, V. R., Dey, D., Leroux, A., Di, J., Urbanek, J., Xiao, L., & Zipunnikov, V. (2017). Re-evaluating the effect of age on physical activity over the lifespan. Preventive medicine, 101, 102–108. https://doi.org/10.1016/j.ypmed.2017.05.030

Salvy, S. J., Roemmich, J. N., Bowker, J. C., Romero, N. D., Stadler, P. J., & Epstein, L. H. (2009). Effect of peers and friends on youth physical activity and motivation to be physically active. Journal of pediatric psychology, 34(2), 217–225. https://doi.org/10.1093/jpepsy/jsn071

Biddle, Stuart J. H. , Atkin, Andrew J. , Cavill, Nick and Foster, Charlie(2011) 'Correlates of physical activity in youth: a review of quantitative systematic reviews', International Review of Sport and Exercise Psychology, 4:1, 25 — 49. URL:http://dx.doi.org/10.1080/1750984X.2010.548528

Shao, T., & Zhou, X. (2023). Correlates of physical activity habits in adolescents: A systematic review. Frontiers in physiology, 14, 1131195. https://doi.org/10.3389/fphys.2023.1131195

Choi, J., Lee, M., Lee, Jk. et al. Correlates associated with participation in physical activity among adults: a systematic review of reviews and update. BMC Public Health 17, 356 (2017). https://doi.org/10.1186/s12889-017-4255-2

Alixe H M Kilgour, Matthew Rutherford, Joanna Higson, Samantha J Meredith, Jessica McNiff, Stephanie Mitchell, Anusan Wijayendran, Stephen E R Lim, Susan D Shenkin, Barriers and motivators to undertaking physical activity in adults over 70—a systematic review of the quantitative literature, Age and Ageing, Volume 53, Issue 4, April 2024, afae080, https://doi.org/10.1093/ageing/afae080

Lee IM, Shiroma EJ, Lobelo F, Puska P, Blair SN, Katzmarzyk PT. Effect of physical inactivity on major non-communicable diseases worldwide: an analysis of burden of disease and life expectancy. Lancet (London, England). 2012, 380(9838):219–229.

Hallal PC, Andersen LB, Bull FC, Guthold R, Haskell W, Ekelund U. Global physical activity levels: surveillance progress, pitfalls, and prospects. Lancet (London, England). 2012, 380(9838):247–257.

Tapia-Serrano, M. Á., López-Gajardo, M. Á., Sánchez-Miguel, P. A., Llanos-Muñoz, R., & Burgueño, R. (2024). Analysis of motivational profiles of physical activity behavior in primary school students: A self-determination theory-based perspective. Personality and Individual Differences, 231, 112837. https://doi.org/10.1016/j.paid.2024.112837

Sebire, S.J., Jago, R., Fox, K.R. et al. Testing a self-determination theory model of children’s physical activity motivation: a cross-sectional study. Int J Behav Nutr Phys Act 10, 111 (2013). https://doi.org/10.1186/1479-5868-10-111

Heredia-León, D. A., Valero-Valenzuela, A., Gómez-Mármol, A., & Manzano-Sánchez, D. (2023). Motivational Profiles in Physical Education: Differences at the Psychosocial, Gender, Age and Extracurricular Sports Practice Levels. Children (Basel, Switzerland), 10(1), 112. https://doi.org/10.3390/children10010112

Geller, K., Renneke, K., Custer, S., & Tigue, G. (2018). Intrinsic and Extrinsic Motives Support Adults' Regular Physical Activity Maintenance. Sports medicine international open, 2(3), E62–E66. https://doi.org/10.1055/a-0620-9137

Brunet, J., & Sabiston, C. M. (2011). Exploring motivation for physical activity across the adult lifespan. Psychology of Sport and Exercise, 12(2), 99–105. https://doi.org/10.1016/j.psychsport.2010.09.006

Oyibo, K., Adaji, I., & Vassileva, J. (2018). Social cognitive determinants of exercise behavior in the context of behavior modeling: a mixed method approach. Digital health, 4, 2055207618811555. https://doi.org/10.1177/2055207618811555

Netz, Yael & Raviv, Shulamith. (2004). Age Differences in Motivational Orientation Toward Physical Activity: An Application of Social—Cognitive Theory. The Journal of psychology. 138. 35-48. 10.3200/JRLP.138.1.35-48. 

Beauchamp, M. R., Crawford, K. L., & Jackson, B. (2019). Social cognitive theory and physical activity: Mechanisms of behavior change, critique, and legacy. Psychology of Sport and Exercise, 42, 110–117. https://doi.org/10.1016/j.psychsport.2018.11.009

de Maio Nascimento, M., Gouveia, É. R., Gouveia, B. R., Marques, A., França, C., Campos, P., Martins, F., García-Mayor, J., & Ihle, A. (2023). Differential Patterns in Motivations for Practicing Sport and Their Effects on Physical Activity Engagement across the Lifespan. Healthcare (Basel, Switzerland), 11(2), 274. https://doi.org/10.3390/healthcare11020274

National, regional, and global trends in insufficient physical activity among adults from 2000 to 2022: a pooled analysis of 507 population-based surveys with 5·7 million participants. Strain, TessaAbdul Raheem, Raheema et al. The Lancet Global Health, Volume 12, Issue 8, e1232 - e1243


Neipp, Mª & Quiles, María & Leon, Eva & Rodríguez-Marín, Jesús. (2013). Theory of Planned Behavior and physical exercise: Differences between people who do regular physical exercise and those who do not. Wulfenia. 20. 324-335. 


Wang, L., & Wang, L. (2015). Using Theory of Planned Behavior to Predict the Physical Activity of Children: Probing Gender Differences. BioMed research international, 2015, 536904. https://doi.org/10.1155/2015/536904

\newpage

# Appendix A - Survey Questions

## Adults

```{r va, fig.align='center',tab.cap="Adult survey questions."}
opts <- options(knitr.kable.NA = '')
kable(list(va),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```


## Youths

```{r vc, fig.align='center',tab.cap="Youth survey questions."}
opts <- options(knitr.kable.NA = '')
kable(list(vc),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

\newpage

# Appendix B - Exercise Types

## Adults

The total minutes of exercise in this dissertation are calculated as the sum of vigorous (original variable: DUR_HVY_CAPPED_SPORTCOUNT_A01) and moderate plus (original variable: DUR_MOD_CAPPED_SPORTCOUNT_A01) activities.

### Vigorous Exercises
Definition
Activity capped: Moderate intensity minutes per week: Sport (count definition), capped at 1680 mins/wk.

* 11 a-side football
* 13 a-side rugby league
* 15 a-side rugby union
* Bootcamp (e.g. drill sergeant military fitness)
* Boxing
* Boxing class (e.g. Boxercise body combat)
* Cardio class (e.g. aerobics step aerobics body attack)
* Circuit or cross training, cross fit, HIT or boot camp
* Circuit training
* Cross fit
* Cross training
* Cycle class (e.g. spinning RPM)
* Cyclo-cross
* Fell or trail running
* Field hockey
* Futsal
* High intensity (e.g.  HIT insanity)
* Hockey
* Indoor cycling - in a class
* Ju-Jitsu
* Karate
* Mountain biking
* Obstacle course (e.g. Tough Mudder Spartan Rat Race)
* Parkour or free running
* Road cycling or racing
* Rugby union
* Running or jogging
* Small sided football
* Squash or racketball
* Taekwondo
* Track and field athletics
* Triathlon (includes aquathlon and duathlon)
* Weightlifting or powerlifting (using a barbell)


### Moderate Exercises
Definition: Activity capped: Moderate intensity minutes per week: Sport (count definition), capped at 1680 mins/wk.

* A session combining several gym or fitness machines or activities
* Aikido
* Badminton
* Baseball or softball
* Basketball
* BMX
* Body weight exercises (e.g. pull ups press ups sit ups)
* Cheerleading
* Chinese martial arts
* Climbing or bouldering wall
* Climbing or mountaineering
* Cross training machine (e.g. Cross trainer SkiErg)
* Cycling for leisure
* Cycling for leisure
* Cycling for Leisure and all other cycling
* Cycling for travel (including commuting)
* Cycling for travel incl commuting
* Dance-based class (e.g. Zumba fitsteps ravercise or body jam)
* Dance-based class (e.g. Zumba, fitsteps, ravercise or body jam) 
* Dressage
* Eventing
* Exercise bike
* Exercise machine
* Football
* Free weights (includes kettlebells and dumb-bells)
* Gymnastics
* Gymnastics or trampolining
* Handball
* Hill and mountain walking, hiking, mountaineering
* Hill or mountain walking or hiking
* Hill or mountain walking or hiking
* Indoor cycling - not in a class
* Judo
* Lacrosse
* Martial arts
* Mountaineering and scrambling
* Netball
* Other exercise machine
* Other football
* Other horse riding
* Resistance weights machines
* Rock climbing or bouldering
* Rounders
* Rounders
* Rowing
* Rowing machine
* Rowing machine
* Rugby league
* Rugby sevens
* Running machine or treadmill
* Schooling
* Show jumping
* Skiing
* Skiing or snowboarding
* Skipping
* Snowboarding
* Step machine
* Surfing, body surfing or body boarding
* Tag or other rugby league
* Tag or other rugby union
* Tennis
* Touch rugby
* Touch rugby league
* Touch rugby league
* Touch rugby union
* Touch rugby union
* Track cycling
* Trampolining
* Treadmill
* Volleyball
* Walking football
* Water based rowing
* Water polo
* Water polo
* Water-based class (e.g. aquaerobics aquafit)
* Weights (did not specify whether free weights or resistance weights)
* Weights-based class (e.g. body pump kettlebell)
* Wrestling

## Youths

Moderate and vigorous activities are encompassed under the same variable (original variable name: mins_modplus_outschool_Week_ALL). 
Deinition: Mins spent in week (moderate plus mins) outside school: All activities

According to Sports England (2024), activities were categorized based on the following definition:

> Moderate activity: This is defined as activity where you raise your heart rate and feel a little out of breath (In academic year 2017-18 (Year 1) pupils were asked whether it made them breathe faster, but since academic year 2019-20 (Year 3) have been asked whether it made them breathe faster than sitting down reading. In Year 2 (18-19), half the children were asked the year 3 version and half were asked the year 1 version across the whole year).

> Vigorous activity: This is defined as when you are out of breath or are sweating - you may not be able to say more than a few words without pausing for breath (pupils were asked whether it made them hot or tired).

More specifically, these acvities were included

* Cycling for fun
* Dancing (include online or TV led e.g. TikTok dances)
* Trampolining (including in a garden, at a trampoline centre, or as part of a club)
* Playing it, tag, chase, sardines or other running game
* Football
* Netball
* Hockey
* Rugby (including tag rugby)
* Touch or tag Rugby
* Contact rugby (rugby union)
* Rugby league (contact)
* Basketball
* Cheerleading
* Running, jogging, cross-country
* Field athletics
* Gym or fitness (fitness/online class e.g., push-ups, sit-ups, yoga, etc or using exercise machines e.g. rowing machine, exercise bike, running machine)
* Judo, karate, taekwondo and other martial arts
* Sports day events
* Boxing


\newpage

# Appendix C - Other tables

```{r adult.summary, fig.align='center',tab.cap="Statistics on Likert-scale responses in adults."}
opts <- options(knitr.kable.NA = '')
kable(list(adult.summary),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

```{r child.summary, fig.align='center',tab.cap="Statistics on Likert-scale responses in youths."}
opts <- options(knitr.kable.NA = '')
kable(list(child.summary),align='c',booktabs = T) %>%
  kable_styling(position = 'center', font_size = 10,
                latex_options = c('striped',
                                  'hold_position'))
```

\newpage

# Appendix D - R Code (Data Cleaning)

```{r data.R, comment = ""}
source("data.R", echo = T, print.eval = F,
       max.deparse.length=Inf, keep.source=T)
```

\newpage

# Appendix E - R Code (SEM)
```{r SEM.R, comment = ""}
source("SEM.R", echo = T, print.eval = F,
       max.deparse.length=Inf, keep.source=T)
```

\newpage

# Appendix F - R Code (LCA)
```{r LCA.R, comment = ""}
source("LCA.R", echo = T, print.eval = F,
       max.deparse.length=Inf, keep.source=T)
```

\newpage

# Appendix G - R Code (Visualization)
```{r Visualization.R, comment = ""}
source("Visualization.R", echo = T, print.eval = F,
       max.deparse.length=Inf, keep.source=T)
```

\newpage

# Appendix H - R code (R Markdown)

```{r get-labels, echo=FALSE, eval=TRUE}
labs = knitr::all_labels()
labs = setdiff(labs, c("setup", "get-labels"))
```

```{r all-code, ref.label=labs, echo=TRUE, eval=FALSE}
```